### 2013-04-28 writing/negatives/thesis_neg_results.tex
initial migration from old structure

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[margin=0.75in]{geometry}
\usepackage{setspace}
\begin{document}
\doublespacing

I tried developing inSAR of the region. I was unable to interpret the results. I'm not sure what they mean. Here are the images. If you have suggestions, I would very much appreciate them. (ADD THE IMAGES along with a similar spatial extent google earth image)

I tried using a global seismic model to get some handle on the upper mantle and lower crustal P and S wave velocities under the basin. I was unaware about how low resolution these models are. It's quite clear that something like the Shargyn Basin is too small to show up. I tried the vanderHilst group model and multiple of the S-wave models.

I tried using crustal thickness models. These are not high enough resolution for something like the Shargyn Basin... I used Gabi Laske's crust 5.1 model.

	Most of this discussion assumes there are no significant faults deforming the Shargyn basin. The surface clearly lacks any noticeable faults. However, there could be blind structures beneath the surface. I do not have the tools to examine this possibility, so I will proceed assuming that no faults deform the Shargyn Basin.

	The absence of active extension still leaves the possibility that the Shargyn Basin is a remnant feature from a time when extension was more prevalent. Erosion rates in Mongolia have been very low for the last 150 Ma \citep{Jolivet2007}. But, given the rate of deformation in recent times, even if the Shargyn Basin were a leftover feature from a previous tectonic regime, we would expect the active faults to be deforming it. Because the faults are not deforming (uplifting) the basin,

\end{document}

### 2013-04-28 writing/paper/main.tex
initial migration from old structure

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[margin=0.75in]{geometry}
\usepackage{setspace}
\begin{document}
\doublespacing

\tableofcontents
\listoffigures
\listoftables

\section{Introduction}
\subsection{Motivation}
	Continental deformation occurs far from plate boundaries. The behavior of intraplate fault systems is poorly understood. Furthermore, complex kinematics can occur at the intersection of multiple fault systems. Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia. The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}. The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history. Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6.5in, height = 4in]{broad.png}
	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes. Important locations or faults are labeled. The study area and location of Figure \ref{regional} is outlined in black. Modified after \citet{Calais2003}}
	\label{broad}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6in, height = 4in]{regional.png}
	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
	\label{regional}
\end{figure}

	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system. In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics. The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays. I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.

\subsection{Mongolian Deformation}
For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high. GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}. The average rate of motion is approximately 5 mm/year. Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century. In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia. The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}. It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}. These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia. The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.

	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision. However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}. Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}. Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}. Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.  Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}. The far-field effects of this model are a major component of the recent stresses in Mongolia.

	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}. Classically, researchers described the EASS as a free-boundary. But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision. However, the potential for far-field stresses from subduction zone rollback is not well constrained. Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins. It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene. A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone. Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6.6in, height = 6in]{asia.png}
	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia. 1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor. BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}. I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia. Gray arrows are added to indicate the direction of motion. The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line. Note the continued left-slip shear all the way to the Pacific coast.}
\end{figure}

	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran. Many of these faults terminate in large thrust systems like the Tian Shan. One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear. (CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan. Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.

	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}. The region is dominated by NW-SE striking right-lateral slip. The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults. Some of these thrusts are isolated. Others are linked by a right-slip faults. In some places, the thrusts are part of a restraining bend. Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.

Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}. North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures. Most of the topography is formed by restraining bends or isolated thrusts. The deformation is diffuse over a 250-350km north-south extent. The range accomodate the eastwards and northeastwards deformation as measured by GPS. At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan). There are no slip estimates for the Bogd Fault System in the north. However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.  Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.

(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems). However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}. This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.

\subsection{Shargyn Basin}

	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system. It is approximately 150km in east-west extent and 100km in north-south extent. A small ephemeral lake is located in the center of the basin. Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited. For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body. It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}. To the east and south, the tectonic regime is less certain. The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there. However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.  The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4.6in, height = 4in]{basins.png}
  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
  \label{basintypes}
\end{figure}

At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
\begin{itemize}
	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure. Shown in Figure \ref{basintypes}C
	\item Compresional stepover basins form similarly to restraining bend basins. However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault. In the compressional case, the basin is essentially a foreland basin in front of the thrust. Arguably, these fault systems are just large restraining bends. Shown in Figure \ref{basintypes}D.
\end{itemize}

	A goal of this paper is to describe the Shargyn Basin in these terms. A summary of the results is in the Structural Overview section. Nearby regions have already been investigated. Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range. The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m. It represents the southeastern-most uplift in the Altai. The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}. A stepover in the right lateral fault has resulted in progressively more thrusting. The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}. To the east, the Dariv Thrust uplifts the Dariv Mountains. \citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust. The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range. Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.

\subsection{Bedrock Geology}
The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt. The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}. The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.

The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age. North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks. The area is dominated by greenschist facies accretionary complex units. In the Dariv Range, arc-related igneous units are prevalent. The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units. To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks. Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}. Small, post-orogenic sedimentary units are also present throughout the region.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
  \caption{Schematic tectonostraigraphic map of Mongolia. The area we studied is located in the outlined box. The Main Mongolian Lineament is the black line passing through our study region. Credit to \citet{Windley2007}}
  \label{bedrock}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{grain.png}
  \caption{Schematic map of the dominant structural grain in western Mongolia. The study area is boxed for reference. Two large, rigid crustal blocks are shaded. The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny. The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}. Modified after \citet{Cunningham2005a}}
  \label{grain}
\end{figure}

The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness. The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}. There is a close correspondence between the major faults and this structural fabric. The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.

\clearpage
\section{Structural Overview}
The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover. The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin. The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts. However, some motion remains on the continuation of the Bogd Fault. The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge. Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range. Figure \ref{schematic} shows a schematic diagram of this structure.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
  \caption{Schematic structural fault map of the Shargyn Basin region. The extent is approximately the same as the extent in Figure \ref{regional}. Description of the structure is in the text. The shaded areas are basins. Unshaded areas are mountainous. The dotted lines represent the dominant foliation strike.}
  \label{schematic}
\end{figure}

	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts. The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault. On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.

\subsection{Remote Sensing}
Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers). I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery. Surface features indicating active motion were identified. Because the region is arid, many important features are clearly visible. For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature. Using such features from Landsat imagery was useful in identifying important features for further investigation in the field. Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.

\subsection{Field Mapping}
During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area). We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}). We mapped at a broad scale, but made careful note of features indicating recent brittle deformation. Mapping was performed on paper with the assistance of satellite imagery and a GPS device. We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators. None of the sample analysis is included in this document.

The results of this mapping are summarized in a fault map in Figure \ref{faultmap}. Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated. Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated. The most notable features of the map were summarized above. The following sections describe in detail the important structural results.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts). The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections. The white boundaries separate different mapping the areas. The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed. The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}. The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
  \label{faultmap}
\end{figure}

\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
\end{figure}

The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT). The cross-section across these two faults is marked on the map as E-E'. Other minor faults are identifiable on the satellite imagery. We did not investigate the minor faults in the field. Both the UKT and LKT strike approximately NNW-SSE. The UKT is the northern fault. On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies. At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present. No outcrop is found on the southwest side of the UKT. At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself. This zone is shown in Figure \ref{faultzones}A. Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$. Dips were approximately similar to the dip of the UKT. Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 6in]{faultzones.png}
  \caption{Photographs of exposed faults. Field workers are used as scale bars. A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'. B) An outcrop of the exposed thrust in cross-section G-G'}
  \ref{faultzones}
\end{figure}

The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT. South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium. Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds. Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium. This marks the fault trace for the LKT. The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?). The older alluvium further northwest also support a complex faulting history.

There is evidence that the motion on the LKT began after the motion on the UKT. In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons. Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed. The uplift was rapid enough that the drainage pattern was "locked in". Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase. Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.

\subsection{Shargyn Fault and Termination Thrust Splays}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
  \caption{Cross sections through the Shargyn Fault and its termination splays.}
  \label{shargynfaultxsecs}
\end{figure}
The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault. Four of the transects focus on this fault. The fault strikes WSW-ENE. I have assumed that the fault is approximately vertical because the motion is strike-slip.

The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}. The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM). However, our interpretation of the location of the thrusts is different from theirs. To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin. Little Dariv Mountain is located on the boundary of these two basins. The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks. On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls. Separating the range is a thrust fault. The fault follows a valley that is clearly visible on satellite imagery. To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards. No other faults were identifiable in the core of Little Dariv Mountain.

On the cross-section, two other faults are indicated on the periphery of the mountain. We did not investigate either of these faults in the field. However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments. There is not much evidence for the northern fault. However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault. To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears. Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.

In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}. To the north, the topography rises steeply towards a relatively flat plateau. The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts. Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here. Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range. It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.

The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}. The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE). It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side. Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}). Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation. There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.  The fault itself passes through a topographic low and does not outcrop. On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation. Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade. The topography and lithology shifts both provide evidence for the presence of a fault at the marked location. The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming. The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
  \label{shargynoffsetstreams}
\end{figure}

To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}. Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE). To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss. To the north of the fault, the topography is lower and transitions into alluvium. The northern unit is an slightly deformed alkali granite (show picture that I took?). Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault. The topographic depression along the fault provides further evidence. To the east, the fault terminates somewhere near the city of Altai. There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.

\subsection{Tonhil Fault and Intersection Wedge Thrusts}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
  \caption{Cross sections through the intersection wedge.}
  \label{intersectionwedgexsecs}
\end{figure}

Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.

At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}. After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south. However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault. There is a noticeable topographic rise bending away form the main Bogd Fault. Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion. In addition, there was a higher local degree of damage to the bedrock. We observed a fracture set at (FILL IN S/D).

At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking. At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'. This has been previously identified on satellite imagery by authors (CITE THEM!). We examined the area. Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise. North of the basin, we found a large granite intrusion. South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks. Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.

The western boundary of the intersection wedge is controlled by the Tonhil Fault. The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault. At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge. At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S. This bend appears to have created another transtensional basin. We did not examine the transtensional basin in the field. However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.

Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side. This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts. For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.

\subsection{Bogd Fault and Termination Thrust Splays}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
  \caption{Cross sections through the Shargyn Fault and its termination splays.}
  \label{bogdfaultxsecs}
\end{figure}

Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault. These faults can be seen in the large scale fault map. Two transects were studied across these thrust splays.

The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}. Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body. A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself. The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery. This is confirmed in the field, though minor folding interferes locally. This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit. Further south along the cross-section, we encounter another fault. The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence. Furthermore, there is another jump in topography at the fault. This fault was never observed. However, the satellite imagery indicates that it dominates the topographic expression west of our fault.

Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'. Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F'). This fault was visible in outcrop. The fault strikes approximately east and dips 15 degrees. The foot wall consists of low grade metamorphosed mudstones. The hanging wall consists of similar clastic metasedimentary rocks plus greenschist. Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge. Kinematic indicators suggest thrust sense of motion. A photograph is shown in Figure \ref{faultzones}B. The topography rises sharply directly south of the location of the fault.

\section{Geophysical Data}
	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology. I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements. I examined the available gravity, magnetic and seismic data for the region.

	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area. Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid. A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly. This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies. The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.

	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160. Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly. Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly. The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement. The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions. \citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km. This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin. Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range. The density anomalies could be linked to the location of past and present faulting. Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
  \caption{Bouguer gravitational anomalies in western Mongolia. The Shargyn Basin is outlined in black. Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges. This is the opposite of what we would expect from isostatic considerations.}
  \label{gravity}
\end{figure}

	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin. I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.  The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}. The Shargyn Basin is at the location of a small positive magnetic anomaly. However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin. This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km. It is also likely that there is significant spatial error near the minimum wavelength. The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths. Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
  \label{magnetic}
\end{figure}

	Earthquake focal mechanisms provide useful information about active fault slip in the region. The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data. \citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia. On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip. In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting. Both possibilities are reasonable given the location of the earthquake.

	Seismic wave speeds underneath the basin could indicate different crustal properties. I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin. If there have been local seismological surveys performed, I am unaware of them.

\section{Formation of the Shargyn Basin}
	It is interesting to ask why the Shargyn Basin exists. The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault. Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts. However, further north, this westward motion continues. The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust. But, why does this stepover form at all? Moreover, why did it form in its current position? What does the answer tells us about the tectonic processes at work in the area? These are all fundamentally the same question: what causes a given fault to form in a specific location? This is a deep and difficult question to answer in the general case. However, for the Shargyn Basin, there is a partial answer.

	The first answer to this questions involves pre-existing weaknesses in the crust. Near the stepover, the Bogd Fault bends southward 20-30 degrees. For most of its length, it closely follows the Main Mongolian Lineament. The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side. The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament. Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area. On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.

	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.  The foliations could be explained by dragging along the fault. With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike. Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge. But, in other locations, such dragging is not visible. The foliations also fit well with a model where faults form along structural weaknesses. Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.

	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement. Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin. Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly. But, it does indicate that something is different. The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block. The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.

	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}. A very steep topographic rise surrounds the north and west margins of the topographically low basin. The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}. The Tarim basin is much larger. The Tarim basin is thought to be a cratonal block, leading to its strength. The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}). The northern border of the Tarim Basin is the Tian Shan. The Tian Shan is especially mechanically weak, leading to the concentration of deformation. Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}. The lack of active faults or topography cutting the basins is an obvious indication of their strength. The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.

	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin. But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins. Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin. We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin. The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault. The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.

	Notably, we have explained a large basin without any extension. Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process. Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005). The low elevations do not require extension. The "Valley of the Lakes" to the north is at similar elevations. The small transtensional basins near the intersection wedge were the only extensional features we found. However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin. Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}. These features are also much smaller than the Shargyn Basin. In addition to these arguments, the low elevations could be due to the positive density anomaly. Denser crust would naturally isostatically adjust to be topographically low.

\section{Conclusion}
	The India-Eurasia collision causes deformation through a broad swath of Asia. In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate. Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns. Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect. Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between. The Shargyn Basin is an anomalously large basin for the region and begs explaining.

	We describe the basin as a combination of two active structures. A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault. A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin. Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.

	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data. Some difference in crustal basement may explain the density anomaly. Such crustal basement may be mechanically strong and difficult to fault. So, this unique block of crust may be controlling the deformation in the area. The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults. Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.

\bibliographystyle{plainnat}
\bibliography{/home/tbent/projects/library.bib}

\end{document}

### 2013-04-28 writing/proposal/proposal.tex
initial migration from old structure

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[margin=0.75in]{geometry}
\begin{document}
\section*{Problem}
	The Altai and Gobi-Altai are two large, actively forming, mountain ranges in central Asia. The Altai are characterized by right-lateral transpressional slip, while the Gobi-Altai are characterized by left-lateral transpressional slip \citep{Cunningham2005a}. The intersection of these two fault systems provides a natural laboratory for studying complex fault behaviors. Furthermore, given the great distance to a plate boundary, the region provides a useful perspective on the far-field stresses of the Indian collision and the East Asian Subduction System.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4.5in, height = 4in]{regional.png}
  \caption{Topographic map of Mongolia with fault overlays and seismic moment tensor solutions for recent earthquakes. The Altai, in the west, and the Gobi-Altai in the center, are labeled. Credit to \citep{Calais2003}}
\end{figure}

	At the intersection of the two fault systems, there is a Connecticut-sized basin called the Shargyn basin. The basin is surrounded by various mountain ranges. To date, no coherent explanation for the basin has been proposed. Because it is in a region dominated by compression and shows no indication of extension \citep{Cunningham2005a}, and is surrounded on all sides by mountain ranges, the formation mechanism of the basin is difficult to understand. For my senior thesis, I intend to use data collected on the active and recent deformation surrounding the Shargyn basin to gain some insight into how the basin formed. It is also possible that the work on the Shargyn basin area would inform the debate about far-field stresses in Eurasia.

\begin{figure}[h!]
	\centering
	\includegraphics[width = 6in, height = 4in]{local.png}
	\caption{Satellite image of the study area, showing notable features. Question marks indicate regions for which we had no reliable data prior to our fieldwork.}
\end{figure}

\section*{Mongolian Deformation}
	For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high. GPS measurements show that ongoing motion is to the north in the Altai, and rapidly rotates eastward in the Gobi-Altai\citep{Calais2003}. Most rates are near 5mm/yr. This motion has resulted in multiple earthquakes with a moment magnitude of 8 or greater in the last century. In July 1905, earthquakes on the Bolnay fault in northern Mongolia had ruptures with magnitudes of 7.9 and 8.4. The December 4th, 1957 earthquake on the Gobi-Altai fault is one of the best exposed surface ruptures in the world of a major strike-slip earthquake \citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}. It has even been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}. But, why do we get crustal deformation in Mongolia when it is so far from a plate boundary?

	The collision between India and Eurasia is generally cited as the source of the driving stresses for the faults forming the Altai and Gobi-Altai. A classic paper by \citep{Tapponnier1982} demonstrated that many of the macro-scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone). Most importantly, this model demonstrated that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.  Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}. The far-field effects of such models are a major component of the recent stresses in Mongolia.

\begin{figure}[h!]
	\centering
	\includegraphics[width = 5in, height = 5in]{asia.png}
	\caption{Tectonic map of Asia. Large arrows show direction of crustal motion. Credit to \citep{Tapponnier1982}}
\end{figure}

	However, theoretical models of crustal deformation in Eurasia suggest that the current crustal motions in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}. Also, there are indications that some of the current crustal motion in Mongolia began before the India-Eurasia collision \citep{Schellart2005}. Both these problems require a different source for the driving stresses behind Mongolian crustal motion.

	Another potential source of stress is rollback on the East Asian Subduction System \citep{Schellart2005}. Subduction zone rollback occurs when the boundary between the subducted plate and the overlying plate moves oceanwards -- towards the subducting plate. Because, the subduction zone "pulls" on the overlying plate, extension occurs in the back-arc region. This is often associated with the formation of small ocean basins in back-arc of a major subduction zone. The potential for far-field stresses from subduction zone roll-back are not well constrained. Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins. A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, then intraplate deformation could occur as far as 3000km from the subduction zone.

\section*{Shargyn Basin Deformation}

	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system. It is a Connecticut-sized basin, bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}. To the east and south, the tectonic regime is less certain. The  Khantaishir Range to east has been the subject of igneous petrology studies concerning the ophiolite located there. However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.  To the south, a wide mountain range (name unknown) bounds the basin.

	Not quite bordering the basin, to the northwest, is the Sutai Range. It is the largest mountain range in the region, with a ice-capped summit above 4000m. A significant amount of structural work has been done on the Sutai Range and the nearby Dariv Basin. The Sutai Range is an active restraining bend in the Altai fault system. A stepover in the right lateral fault has resulted in progressively more thrusting. The faults bounding the range exhibit a flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}. To the east, the Dariv Thrust uplifts the Dariv Mountains. \citep{Howard2006} suggest that the Dariv Basin may be a "piggyback" basin, riding on top of the block thrust up by the Dariv Thrust. South of this, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains.

	Because of its tectonic setting and the predominance of transpressional slip, the Shargyn Basin can be classified as a intracontinental strike-slip basin \citep{Busby1995}. Transpressional basins normally result from flexural subsidence. Sufficient thrusting results in enough added tectonic loading to cause subsidence in adjacent areas, creating a basin. Such thrusting might be expected in a tight restraining bend. The deposition zone is analogous to a very small foreland basin. Often these are paired on opposite sides of a restraining bend. Sometimes, transpressional basins can result from differential extensional stresses due to moving past the curved fault. This can result in small down dropped wedges at restraining bends. The adjacent uplift due to the restraining bend provides a sediment source. Other types of basins can form where there is significant thrusting. As described above, piggyback basins sometimes form on top of young or minor thrust sheets. Although current motion is transpressional, it is possible past motion was transtensional. It is also possible that a releasing bend in a primarily transpressional setting would result in a transtensional pull apart basin.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 3in, height = 4in]{strikeslipbasins.png}
  \caption{The two major types of strike slip basins. The gray dot-filled area represents the basin. Credit to \citep{Busby1995}}
\end{figure}

	Mongolia lies in the distant foreland of the India-Eurasia collision. Because of this setting, it might make sense to compare the Shargyn Basin to intermontane basins, like those found in the Rocky Mountains \citep{Busby1995}. However, the scales and motions are totally different. The Rocky Mountains are massive basement cored uplifts with little strike slip motion, whereas the Altai and Gobi-Altai have significantly less thrusting and show majority strike-slip motion.

	Because the Shargyn basin lies near the intersection of multiple restraining bends in two fault systems with opposite motions, is adjacent to a few large mountain ranges, and has had a complex and varied history, it is probable that some complex combination of strike-slip related factors are involved in its formation.

\section*{Completed Work}
	During the summer of 2012, Claire Bucholz, an MIT EAPS graduate student, and I performed two months of fieldwork in western Mongolia in the bounding mountain ranges surrounding the Shargyn Basin. We performed 5-15 mile cross-sections through areas with important faults, as previously identified by satellite image. We mapped at a broad scale, but made careful note of features indicating recent brittle deformation. We took samples at locations that might provide useful kinematic indicators. We also took samples crossing some of the major faults, possibly allowing us to determine the uplift rate using thermochronology.

	Over the last couple months, I have taken much of the mapped data from the summer, cleaned it, and digitized it. I have added mapping work from other sources. I've also worked on cross-sections through the area.
\section*{Plans} 
	 Moving forward, I will to continue work on the map. I will add more mapping data from neighboring areas I will also create geologic cross-sections across some of the more interesting areas that we studied. To start, I plan to make a cross-section in the east-west direction and a cross-section in the north-south direction. These will provide a launching point for discussions about the kinematics of the basin. A few of the samples collected contain useful kinematic indicators, which, viewed under thin section could test hypothesis about recent fault motion. Both of these tasks will hopefully illuminate some of the recent crustal motions in the area. Based on the data collected, I will develop a kinematic explanation for how the Shargyn basin formed. Given that the Shargyn basin appears to be unique and counterintuitive, I will relate my explanation to the literature on basin formation. Learning about general basin formation concepts, and, more specifically, about how specific basins have formed, will help to understand the Shargyn Basin.

	If enough information is available, I will extrapolate the geologic history back through the Cenozoic. Using the data we collected and satellite imagery, I will look for indications that some faults were active and are no longer active. I will also look for faults that may have changed their direction of slip. I've been communicating with Devin McPhillips, a postdoc at The University of Vermont, about geomorphological records of fault motion and fault slip. Extrapolating into the recent geologic past may provide some insight into when Indian collision stresses began influencing the region and whether, prior to that, East Asian rollback-related stresses were important in the region. More generally, it will help to determine the time-evolution of the stress field in Central Asia.

	If time constraints allow for it, I would like to determine the active deformation rates across some of the faults in the region. This would be possible using interferometric synthetic aperture radar (inSAR). This is a technique that determines the change in phase of a radar wave between two similar satellite orbits. By correcting for topography, the orbits of the satellites, and a number of other variables, ground motions can be determined to the millimeter scale. If the satellite orbits are years apart, this can resolve the rate of crustal motion. Such measurements of current ground motions could help determine both which faults remain active, and the partitioning of motion across different faults with similar strike.
\section*{Potential Complications}
	As always, there is a significant possibility that I won't be able to finish everything I set out to do. It is good that there are discrete pieces to the project. With only a complete map and a few cross sections, I would still be able to provide valuable input on how the Shargyn basin formed. Because there are nearly infinite options for continuing the research, I intend to keep a relatively simple timeline. I will continue working on the research itself through March. By that point, whatever data I have will be compiled into the final thesis paper. I plan to have a draft complete by late April.

	It is probable that I won't have enough information to extrapolate very far into the geologic past. Hopefully, some of the satellite imagery analysis will be able to supplement this. However, if I'm limited to making interpretations about the active formation of the Shargyn basin, that will be still be valuable.

	I am confident that I will be able to develop a coherent and consistent kinematic description of the Shargyn basin. However, if something goes totally awry, I will still be able to fall back on a large amount of mapping work that is useful regardless. On the other hand, if I develop a kinematic description but I cannot identify any previously research basins that have similar kinematics to the Shargyn basin, then I will need to have stronger evidence to support my argument.

\bibliographystyle{plainnat}
\bibliography{/home/tbent/projects/library.bib}

\end{document}

### 2013-04-28 writing/paper/grabimages.py

@@ -1,0 +1,36 @@
+#!/usr/bin/python
+import os
+import subprocess
+def call_command(command):
+    process = subprocess.Popen(command.split(' '),
+                               stdout=subprocess.PIPE,
+                               stderr=subprocess.PIPE)
+    return process.communicate()
+
+call_command('../shargynxsecs/composites/shargynfault/make_composite.py')
+call_command('../shargynxsecs/composites/bogdfault/make_composite.py')
+call_command('../shargynxsecs/composites/intersectionwedge/make_composite.py')
+call_command('../shargynxsecs/composites/khantaishir/make_composite.py')
+
+files = dict()
+files['./shargynfault.png'] = '../shargynxsecs/composites/shargynfault/final.png'
+files['./bogdfaultxsecs.png'] = '../shargynxsecs/composites/bogdfault/final.png'
+files['./intersectionwedgexsecs.png'] = '../shargynxsecs/composites/intersectionwedge/final.png'
+files['./khantaishir.png'] = '../shargynxsecs/composites/khantaishir/final.png'
+files['./gravity.ps.png'] = '../gravity/gravity.ps.png'
+files['./magnetic.ps.png'] = '../magnetic/magnetic.ps.png'
+files['./broad.png'] = '../broadfiguremine/broad.png'
+files['./regional.png'] = '../regionalfigure/regional.png'
+files['./basins.png'] = '../basintypesfigure/basins.png'
+files['./asia.png'] = '../asiafigure/asia.png'
+files['./bedrock.png'] = '../bedrockfigure/bedrockfigure.png'
+files['./grain.png'] = '../grainfigure/grain.png'
+files['./faultzones.png'] = '../faultscollage/faultscollage.png'
+files['./schematic.png'] = '../schematicfaults/schematic.png'
+files['./faultmap.png'] = '../maps/markedupfaultmap.png'
+
+for key,value in files.iteritems():
+	cmd = 'cp ' + value + ' ' + key
+	print cmd
+	call_command(cmd)
+

### 2013-05-02 writing/paper/grabimages.py
worked on presentation and redid some figures, moved proposals into archive

@@ -12,6 +12,7 @@ call_command('../shargynxsecs/composites/bogdfault/make_composite.py')
 call_command('../shargynxsecs/composites/intersectionwedge/make_composite.py')
 call_command('../shargynxsecs/composites/khantaishir/make_composite.py')
 
+
 files = dict()
 files['./shargynfault.png'] = '../shargynxsecs/composites/shargynfault/final.png'
 files['./bogdfaultxsecs.png'] = '../shargynxsecs/composites/bogdfault/final.png'

### 2013-05-04 writing/paper/abstract.tex
reformatted thesis, incorporating lots of oli's edits

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. I examine the fault kinematics surrouding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays. The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}

### 2013-05-04 writing/paper/background.tex
reformatted thesis, incorporating lots of oli's edits

--- 2013-04-28 writing/paper/main.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{setspace}
-\begin{document}
-\doublespacing
+\section{Background on Mongolian Active Deformation}
+	Understanding the tectonic context of the Shargyn Basin in western Mongolia and Asia is critical to interpreting the active deformation.
 
-\tableofcontents
-\listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
-The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
-The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history.
-Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
-	\label{regional}
-\end{figure}
-
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
+\subsection{Mongolian Deformation in Context -- Asia}
 
@@
 Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
+The zone with spaced right-slip faults are surrounded by a dashed line.}
+	\label{asia}
 \end{figure}
 
+	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}(cite more).
+Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models.
+The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citet{Tapponier1982}\citep{Peltzer1996}.
+This hypthesis suggests that most motion has been accomodated by large strike-slip and thrust faults.
+The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993}\citep{England1997}.
+In these models, the important effects continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.
+
+	Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
+The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponier1982}\citep{Peltzer1996}\citep{Houseman1993}.
+But, because of its orientation, active EASS rollback would exert stresses in a similar orientation to the Indo-Eurasian collision.
+However, the potential for far-field stresses from subduction zone rollback is not well constrained.
+Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
+Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995}\citep{Schellart2005a}.
+A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
+Additional constraints on the active and past deformation in western Mongolia could influence these broad theories of Asian deformation.
+
+	 The intensity of intracontinental deformation in Mongolia is emphasized by measured GPS motions and three of the largest strike-slip earthquakes ever.
+For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
+GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
+The average rate of motion is approximately 5 mm/year relative to a stable Eurasia as defined by \citet{Calais2003}.
+Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
+In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
+The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
+The Bogd Earthquake has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
+The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
+
 	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
+The area is marked in the northwestern corner of Figure \ref{asia}.
 Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
+One explanation invokes bookshelf-like rotation along a long left-lateral shear zone creates the right-lateral faulting.
+\citep{Davy1998} \citep{Bayasgalan2005a}  This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
+Possibly, the Altai and Gobi-Altai are similar to an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width = 6.5in, height = 4in]{broad.png}
+	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
+Important locations or faults are labeled.
+The study area and location of Figure \ref{regional} is outlined in black.
+The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important.
+Modified after \citet{Calais2003}}
+	\label{broad}
+\end{figure}
 
@@
 The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
 Others are linked by a right-slip faults.
 In some places, the thrusts are part of a restraining bend.
+Major faults, GPS motions, and earthquake moment tensors are shown in Figure \ref{broad}.
 Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
@@
 North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
 The deformation is diffuse over a 250-350km north-south extent.
 The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
+At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System \citep{Cunningham2003a}.
 There are no slip estimates for the Bogd Fault System in the north.
 However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
+Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.
 
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
+	Previous researchers have suggested that the Hangay Dome in central Mongolia (Figure \ref{broad}) is a rigid block causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems) \citep{Cunningham2005a}.
+However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome.
+This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south, and may not be as rigid as previously thought\citep{Walker2006}\citep{Walker2007}\citep{Walker2008}.
 
@@
 A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
+The basin margin varies between 1500m and 2000m, while the center is at an elevation of 963m.
+The basin is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
 To the east and south, the tectonic regime is less certain.
@@
 The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width = 6in, height = 4in]{regional.png}
+	\caption{Shaded relief map showing notable features in the Shargyn Basin region.
+Many of these faults are studied in this paper.
+GPS velocity magnitude for the Altai site is 4.4mm/yr.
+Faults are shown.
+See Figure \ref{faultmap} for the direction of slip.}
+	\label{regional}
+\end{figure}
 
@@
 
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
+Many different types of basins form in strike-slip and compressional settings.
+Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
+Some of the settings we might expect:
 \begin{itemize}
@@
 Shown in Figure \ref{basintypes}D.
+	\item An intersection wedge can form at a conjugate strike-slip intersection.
+Because strike-slip motion on both faults must be transferred into uplift, an uplifted wedge of material forms between the two faults.
 \end{itemize}
@@
 The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
+The shaded Junggar Block is mechanically rigid.
+The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
 The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
+It is not shaded because it is not shown to be rigid like the Junggar Block.
 Modified after \citet{Cunningham2005a}}
@@
 \clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
-\section{Geophysical Data}
-	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
-
-	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area.
-Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid.
-A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly.
-This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies.
-The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.
-
-	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160.
-Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly.
-Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly.
-The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement.
-The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions.
-\citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km.
-This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
-The density anomalies could be linked to the location of past and present faulting.
-Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
-  \caption{Bouguer gravitational anomalies in western Mongolia.
-The Shargyn Basin is outlined in black.
-Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges.
-This is the opposite of what we would expect from isostatic considerations.}
-  \label{gravity}
-\end{figure}
-
-	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
-The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
-The Shargyn Basin is at the location of a small positive magnetic anomaly.
-However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin.
-This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km.
-It is also likely that there is significant spatial error near the minimum wavelength.
-The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths.
-Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
-  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
-  \label{magnetic}
-\end{figure}
-
-	Earthquake focal mechanisms provide useful information about active fault slip in the region.
-The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
-\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
-In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-	Seismic wave speeds underneath the basin could indicate different crustal properties.
-I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin.
-If there have been local seismological surveys performed, I am unaware of them.
-
-\section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
-
-	The first answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
-On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
-
-	Notably, we have explained a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.
-
-\section{Conclusion}
-	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
-Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
-Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect.
-Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between.
-The Shargyn Basin is an anomalously large basin for the region and begs explaining.
-
-	We describe the basin as a combination of two active structures.
-A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault.
-A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin.
-Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
-\bibliographystyle{plainnat}
-\bibliography{/home/tbent/projects/library.bib}
-
-\end{document}

### 2013-05-04 writing/paper/basin.tex
reformatted thesis, incorporating lots of oli's edits

--- 2013-04-28 writing/paper/main.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{setspace}
-\begin{document}
-\doublespacing
-
-\tableofcontents
-\listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
-The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
-The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history.
-Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
-	\label{regional}
-\end{figure}
-
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
-	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
-1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor.
-BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
-\end{figure}
-
-	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
-Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
-
-	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}.
-The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
-Others are linked by a right-slip faults.
-In some places, the thrusts are part of a restraining bend.
-Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
-
-Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}.
-North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
-The deformation is diffuse over a 250-350km north-south extent.
-The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
-There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
-
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
-
-\subsection{Shargyn Basin}
-
-	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
-A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
-To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
-  \label{basintypes}
-\end{figure}
-
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-\begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-\end{itemize}
-
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
-Nearby regions have already been investigated.
-Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range.
-The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m.
-It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
-A stepover in the right lateral fault has resulted in progressively more thrusting.
-The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
-The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
-
-\subsection{Bedrock Geology}
-The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt.
-The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}.
-The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.
-
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
-North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
-The area is dominated by greenschist facies accretionary complex units.
-In the Dariv Range, arc-related igneous units are prevalent.
-The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units.
-To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks.
-Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
-The area we studied is located in the outlined box.
-The Main Mongolian Lineament is the black line passing through our study region.
-Credit to \citet{Windley2007}}
-  \label{bedrock}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{grain.png}
-  \caption{Schematic map of the dominant structural grain in western Mongolia.
-The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-Modified after \citet{Cunningham2005a}}
-  \label{grain}
-\end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
-
-\clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
-\section{Geophysical Data}
-	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
-
-	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area.
-Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid.
-A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly.
-This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies.
-The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.
-
-	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160.
-Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly.
-Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly.
-The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement.
-The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions.
-\citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km.
-This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
-The density anomalies could be linked to the location of past and present faulting.
-Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
-  \caption{Bouguer gravitational anomalies in western Mongolia.
-The Shargyn Basin is outlined in black.
-Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges.
-This is the opposite of what we would expect from isostatic considerations.}
-  \label{gravity}
-\end{figure}
-
-	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
-The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
-The Shargyn Basin is at the location of a small positive magnetic anomaly.
-However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin.
-This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km.
-It is also likely that there is significant spatial error near the minimum wavelength.
-The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths.
-Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
-  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
-  \label{magnetic}
-\end{figure}
-
-	Earthquake focal mechanisms provide useful information about active fault slip in the region.
-The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
-\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
-In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-	Seismic wave speeds underneath the basin could indicate different crustal properties.
-I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin.
-If there have been local seismological surveys performed, I am unaware of them.
-
 \section{Formation of the Shargyn Basin}
@@
 
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
+	As has been discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
 But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
 Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
+We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure, the Main Mongolian Lineament, in the area does not intersect the Shargyn Basin.
 The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
+The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a regionally distinct past tectonic environment in the Shargyn Basin.
 
@@
 Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
+Previous researchers have stated that the region has no normal faulting or transtensional faulting \citep{Cunningham2005a}.
 The low elevations do not require extension.
@@
 Denser crust would naturally isostatically adjust to be topographically low.
-
-\section{Conclusion}
-	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
-Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
-Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect.
-Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between.
-The Shargyn Basin is an anomalously large basin for the region and begs explaining.
-
-	We describe the basin as a combination of two active structures.
-A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault.
-A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin.
-Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
-\bibliographystyle{plainnat}
-\bibliography{/home/tbent/projects/library.bib}
-
-\end{document}

### 2013-05-04 writing/paper/conclusion.tex
reformatted thesis, incorporating lots of oli's edits

--- 2013-04-28 writing/paper/main.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{setspace}
-\begin{document}
-\doublespacing
-
-\tableofcontents
-\listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
-The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
-The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history.
-Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
-	\label{regional}
-\end{figure}
-
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
-	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
-1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor.
-BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
-\end{figure}
-
-	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
-Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
-
-	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}.
-The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
-Others are linked by a right-slip faults.
-In some places, the thrusts are part of a restraining bend.
-Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
-
-Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}.
-North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
-The deformation is diffuse over a 250-350km north-south extent.
-The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
-There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
-
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
-
-\subsection{Shargyn Basin}
-
-	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
-A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
-To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
-  \label{basintypes}
-\end{figure}
-
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-\begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-\end{itemize}
-
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
-Nearby regions have already been investigated.
-Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range.
-The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m.
-It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
-A stepover in the right lateral fault has resulted in progressively more thrusting.
-The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
-The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
-
-\subsection{Bedrock Geology}
-The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt.
-The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}.
-The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.
-
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
-North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
-The area is dominated by greenschist facies accretionary complex units.
-In the Dariv Range, arc-related igneous units are prevalent.
-The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units.
-To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks.
-Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
-The area we studied is located in the outlined box.
-The Main Mongolian Lineament is the black line passing through our study region.
-Credit to \citet{Windley2007}}
-  \label{bedrock}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{grain.png}
-  \caption{Schematic map of the dominant structural grain in western Mongolia.
-The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-Modified after \citet{Cunningham2005a}}
-  \label{grain}
-\end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
-
-\clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
-\section{Geophysical Data}
-	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
-
-	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area.
-Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid.
-A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly.
-This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies.
-The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.
-
-	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160.
-Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly.
-Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly.
-The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement.
-The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions.
-\citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km.
-This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
-The density anomalies could be linked to the location of past and present faulting.
-Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
-  \caption{Bouguer gravitational anomalies in western Mongolia.
-The Shargyn Basin is outlined in black.
-Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges.
-This is the opposite of what we would expect from isostatic considerations.}
-  \label{gravity}
-\end{figure}
-
-	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
-The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
-The Shargyn Basin is at the location of a small positive magnetic anomaly.
-However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin.
-This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km.
-It is also likely that there is significant spatial error near the minimum wavelength.
-The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths.
-Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
-  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
-  \label{magnetic}
-\end{figure}
-
-	Earthquake focal mechanisms provide useful information about active fault slip in the region.
-The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
-\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
-In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-	Seismic wave speeds underneath the basin could indicate different crustal properties.
-I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin.
-If there have been local seismological surveys performed, I am unaware of them.
-
-\section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
-
-	The first answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
-On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
-
-	Notably, we have explained a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.
-
 \section{Conclusion}
@@
 Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
-\bibliographystyle{plainnat}
-\bibliography{/home/tbent/projects/library.bib}
-
-\end{document}

### 2013-05-04 writing/paper/geophysical.tex
reformatted thesis, incorporating lots of oli's edits

--- 2013-04-28 writing/paper/main.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{setspace}
-\begin{document}
-\doublespacing
-
-\tableofcontents
-\listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
-The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
-The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history.
-Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
-	\label{regional}
-\end{figure}
-
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
-	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
-1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor.
-BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
-\end{figure}
-
-	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
-Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
-
-	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}.
-The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
-Others are linked by a right-slip faults.
-In some places, the thrusts are part of a restraining bend.
-Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
-
-Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}.
-North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
-The deformation is diffuse over a 250-350km north-south extent.
-The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
-There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
-
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
-
-\subsection{Shargyn Basin}
-
-	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
-A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
-To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
-  \label{basintypes}
-\end{figure}
-
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-\begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-\end{itemize}
-
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
-Nearby regions have already been investigated.
-Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range.
-The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m.
-It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
-A stepover in the right lateral fault has resulted in progressively more thrusting.
-The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
-The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
-
-\subsection{Bedrock Geology}
-The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt.
-The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}.
-The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.
-
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
-North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
-The area is dominated by greenschist facies accretionary complex units.
-In the Dariv Range, arc-related igneous units are prevalent.
-The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units.
-To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks.
-Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
-The area we studied is located in the outlined box.
-The Main Mongolian Lineament is the black line passing through our study region.
-Credit to \citet{Windley2007}}
-  \label{bedrock}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{grain.png}
-  \caption{Schematic map of the dominant structural grain in western Mongolia.
-The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-Modified after \citet{Cunningham2005a}}
-  \label{grain}
-\end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
-
-\clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
 \section{Geophysical Data}
@@
 If there have been local seismological surveys performed, I am unaware of them.
-
-\section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
-
-	The first answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
-On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
-
-	Notably, we have explained a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.
-
-\section{Conclusion}
-	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
-Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
-Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect.
-Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between.
-The Shargyn Basin is an anomalously large basin for the region and begs explaining.
-
-	We describe the basin as a combination of two active structures.
-A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault.
-A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin.
-Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
-\bibliographystyle{plainnat}
-\bibliography{/home/tbent/projects/library.bib}
-
-\end{document}

### 2013-05-04 writing/paper/intro.tex
reformatted thesis, incorporating lots of oli's edits

--- 2013-04-28 writing/paper/main.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
-\usepackage{setspace}
-\begin{document}
-\doublespacing
+\section{Motivation}
+	Continental deformation can occur far from plate boundaries.
+The behavior of intraplate fault systems is poorly understood.
+Furthermore, where multiple fault systems intersect, complex kinematics can occur.
+Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
 
-\tableofcontents
-\listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
+	In this paper, I examine the intersection of the Altai Range and the Gobi-Altai Range.
 The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
@@
 The study area and location of Figure \ref{regional} is outlined in black.
+The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important.
 Modified after \citet{Calais2003}}
@@
 	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
+	\caption{Shaded relief map showing notable features in the Shargyn Basin region.
+Many of these faults are studied in this paper.
+GPS velocity magnitude for the Altai site is 4.4mm/yr.
+Faults are shown.
+See Figure \ref{faultmap} for the direction of slip.}
 	\label{regional}
@@
 	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
-	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
-1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor.
-BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
-\end{figure}
-
-	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
-Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
-
-	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}.
-The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
-Others are linked by a right-slip faults.
-In some places, the thrusts are part of a restraining bend.
-Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
-
-Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}.
-North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
-The deformation is diffuse over a 250-350km north-south extent.
-The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
-There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
-
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
-
-\subsection{Shargyn Basin}
-
-	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
-A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
-To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
-  \label{basintypes}
-\end{figure}
-
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-\begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-\end{itemize}
-
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
-Nearby regions have already been investigated.
-Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range.
-The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m.
-It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
-A stepover in the right lateral fault has resulted in progressively more thrusting.
-The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
-The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
-
-\subsection{Bedrock Geology}
-The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt.
-The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}.
-The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.
-
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
-North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
-The area is dominated by greenschist facies accretionary complex units.
-In the Dariv Range, arc-related igneous units are prevalent.
-The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units.
-To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks.
-Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
-The area we studied is located in the outlined box.
-The Main Mongolian Lineament is the black line passing through our study region.
-Credit to \citet{Windley2007}}
-  \label{bedrock}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{grain.png}
-  \caption{Schematic map of the dominant structural grain in western Mongolia.
-The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-Modified after \citet{Cunningham2005a}}
-  \label{grain}
-\end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
-
-\clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
-\section{Geophysical Data}
-	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
-
-	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area.
-Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid.
-A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly.
-This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies.
-The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.
-
-	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160.
-Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly.
-Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly.
-The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement.
-The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions.
-\citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km.
-This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
-The density anomalies could be linked to the location of past and present faulting.
-Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
-  \caption{Bouguer gravitational anomalies in western Mongolia.
-The Shargyn Basin is outlined in black.
-Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges.
-This is the opposite of what we would expect from isostatic considerations.}
-  \label{gravity}
-\end{figure}
-
-	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
-The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
-The Shargyn Basin is at the location of a small positive magnetic anomaly.
-However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin.
-This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km.
-It is also likely that there is significant spatial error near the minimum wavelength.
-The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths.
-Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
-  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
-  \label{magnetic}
-\end{figure}
-
-	Earthquake focal mechanisms provide useful information about active fault slip in the region.
-The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
-\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
-In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-	Seismic wave speeds underneath the basin could indicate different crustal properties.
-I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin.
-If there have been local seismological surveys performed, I am unaware of them.
-
-\section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
-
-	The first answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
-On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
-
-	Notably, we have explained a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.
-
-\section{Conclusion}
-	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
-Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
-Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect.
-Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between.
-The Shargyn Basin is an anomalously large basin for the region and begs explaining.
-
-	We describe the basin as a combination of two active structures.
-A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault.
-A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin.
-Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
-\bibliographystyle{plainnat}
-\bibliography{/home/tbent/projects/library.bib}
-
-\end{document}
+Here, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
+The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
+I also use gravity data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.

### 2013-05-04 writing/paper/main.tex
reformatted thesis, incorporating lots of oli's edits

@@
-\documentclass[10pt,a4paper]{article}
+\documentclass[12pt, doublepage, vi]{mitthesis}
 \usepackage[latin1]{inputenc}
@@
 \usepackage{amssymb}
-\usepackage{graphicx}
+\usepackage[demo]{graphicx}
 \usepackage{natbib}
@@
 \usepackage{setspace}
+\renewcommand\thechapter{}
+\renewcommand\thesection{\arabic{section}}
+\renewcommand\thesubsection{\thesection.\arabic{subsection}}
 \begin{document}
-\doublespacing
+\title{Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.}
+\author{Thomas Ben Thompson}
+\department{Department of Earth, Atmospheric and Planetary Science}
+\degree{Bachelor of Science in Earth, Atmospheric and Planetary Sciences}
+\degreemonth{June}
+\degreeyear{2013}
+\copyrightnoticetext{\copyright~ 2013 Thomas Ben Thompson \newline
+Anyone is permitted to reproduce and distribute copies of thesis so long as they give me credit.}
+\thesisdate{May 10, 2013}
+\supervisor{Oliver Jagoutz}{Assistant Professor}
+\chairman{Samuel Bowring}{Chair, Committee on Undergraduate Program}
+\maketitle
+\include{abstract}
+
+\section*{Acknowledgments}
+Primarily, I thank Claire Bucholz (EAPS PhD candidate) for wandering the Mongolian steppes and mountains and teaching me endlessly about geology and EAPS Professor Oliver Jagoutz for supervising and helping me through the whole thesis process.
+I thank the UROP office and the EAPS department for supporting my field work in Mongolia last summer.
 
@@
 \listoffigures
-\listoftables
-
-\section{Introduction}
-\subsection{Motivation}
-	Continental deformation occurs far from plate boundaries.
-The behavior of intraplate fault systems is poorly understood.
-Furthermore, complex kinematics can occur at the intersection of multiple fault systems.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
-The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
-The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history.
-Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region. GPS velocity magnitude for the Altai site is 4.4mm/yr. Faults are shown. See Figure \ref{faultmap} for slip.}
-	\label{regional}
-\end{figure}
-
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-In this report, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics can be described as a combination of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use geophysical data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
-
-\subsection{Mongolian Deformation}
-For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
-The average rate of motion is approximately 5 mm/year.
-Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century.
-In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
-It has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
-These three earthquakes are some of the largest strike-slip earthquakes ever, emphasizing the intensity of intracontinental deformation in Mongolia.
-The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.
-
-	Active deformation in Mongolia is mostly a result of the Indian-Eurasian collision.
-However, theoretical models of crustal deformation in Eurasia suggest that the deformation rates in Mongolia are too large to be entirely explained by the India-Eurasia collision \citep{Calais2002a}.
-Also, some of the deformation in Central and Eastern Asia may have begun before the India-Eurasia collision \citep{Schellart2005}.
-Nonetheless, many of the large scale features of Central and Eastern Asia can be replicated by a rigid impactor (India) hitting a plasticine block (Eurasia) with a free "eastern" surface (the East Asian Subduction Zone)\citep{Tapponnier1982}.
-Most importantly, this model showed that the collision of India would cause blocks of Eurasia to "extrude" to east, into the East Asian Subduction Zone.
-Many of the large-scale features predicted by this model have been confirmed \citep{Schellart2005}\cite{Yin2010}.
-The far-field effects of this model are a major component of the recent stresses in Mongolia.
-
-	 Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-Classically, researchers described the EASS as a free-boundary.
-But, because of its orientation, active East Asian Subduction System rollback would exert similar stresses to the Indian-Eurasian collision.
-However, the potential for far-field stresses from subduction zone rollback is not well constrained.
-Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-It is suggested that much of the East Asian Subduction zone was experiencing rollback during the late Cretaceous until the Mid-Miocene.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
-Additional constraints on the active and past deformation in Mongolia could influence this rollback theory.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
-	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
-1, 2 and 3 are major thrust, strike-slip and normal fault, respectively; 4, 5 and 6 are convergent (filled triangles are for active and open triangles are for inactive boundary), transform and divergent plate boundary, respectively; 7 is land; 8 is sea, with dark grey being basin or ocean floor and light grey being continental shelf or morphological high on basin or ocean floor.
-BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The left-slip shear zone with accompanying right-slip bookshelf faults is surrounded by a dashed line.
-Note the continued left-slip shear all the way to the Pacific coast.}
-\end{figure}
-
-	On the broadest scale, the Altai and Gobi-Altai fault systems may be related to a sequence of regularly-spaced, NW-SE striking right-lateral strike-slip faults \citep{Yin2010} located in a belt from west Mongolia through Iran.
-Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation of these right-lateral faults explain them as accomodating bookshelf-like rotation in a massive 3500km long  to accomodate the left-lateral shear.
-(CITE DELVAUX) \citep{Bayasgalan2005a} This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
-Perhaps the Altai and Gobi-Altai represent an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
-
-	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}.
-The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
-Some of these thrusts are isolated.
-Others are linked by a right-slip faults.
-In some places, the thrusts are part of a restraining bend.
-Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
-
-Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}.
-North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-Most of the topography is formed by restraining bends or isolated thrusts.
-The deformation is diffuse over a 250-350km north-south extent.
-The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (CITE Cunningham 2003 about Tian Shan).
-There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
-Earthquake moment tensors show primarily left-lateral strike-slip in the Gobi-Altai.
-
-(CLEAN UP THIS PARAGRAPH, it's ugly) Previously, it was thought that the Hangay Dome, a large domal uplift in central western Mongolia, behaved as an indenter causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems).
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
-This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south.
-
-\subsection{Shargyn Basin}
-
-	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
-A small ephemeral lake is located in the center of the basin.
-Broad, varying, color bands on satellite imagery indicate that much of the basin sediment is recently deposited.
-For example, on the south side of the Dariv Range, there is a broad dark gray band of sediment deriving from a serpentinite body.
-It is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
-To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
-  \label{basintypes}
-\end{figure}
-
-At least four different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-\begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-\end{itemize}
-
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
-Nearby regions have already been investigated.
-Not quite bordering the Shargyn Basin, to the northwest, is the Sutai Range.
-The Sutai Range is the largest mountain range in the region, with an ice-capped summit above 4000m.
-It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
-A stepover in the right lateral fault has resulted in progressively more thrusting.
-The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
-The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
-
-\subsection{Bedrock Geology}
-The bedrock in western Mongolia formed as part of the Central Asian Orogenic Belt.
-The Central Asian Orogenic Belt formed from 1000 Ma to 250 Ma as an agglomeration of island arcs, accretionary wedges, ophiolites and microcontinents in a complex multiphase development similar to what is seen today in southeast Asia \citep{Windley2007}.
-The broad tectonostratigraphic provinces in Mongolia are shown in Figure \ref{bedrock}.
-
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
-North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
-The area is dominated by greenschist facies accretionary complex units.
-In the Dariv Range, arc-related igneous units are prevalent.
-The northern end of the Dariv Range and the bulk of the Khantaishir Range consist of ophiolite-related mafic and ultramafic units.
-To the south of the Main Mongolian Lineament, the units are primarily Silurian and Carboniferous arc-related metamorphic and igneous rocks.
-Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
-The area we studied is located in the outlined box.
-The Main Mongolian Lineament is the black line passing through our study region.
-Credit to \citet{Windley2007}}
-  \label{bedrock}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{grain.png}
-  \caption{Schematic map of the dominant structural grain in western Mongolia.
-The study area is boxed for reference.
-Two large, rigid crustal blocks are shaded.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-Modified after \citet{Cunningham2005a}}
-  \label{grain}
-\end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
-
 \clearpage
-\section{Structural Overview}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust, but the Shargyn Fault separates the low-lying basin from the adjacent mountain range, creating a "strike-slip foreland" basin.
-
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region (significant time was spent on other research projects in the same area).
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-We also took samples at locations across some of the major faults and in locations that might provide useful kinematic indicators.
-None of the sample analysis is included in this document.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-The most notable features of the map were summarized above.
-The following sections describe in detail the important structural results.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence grading from conglomerate into other clastic facies.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, we observed a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Dips were approximately similar to the dip of the UKT.
-Tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and flow basalts, with minor sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northwest of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure was also identified by previous researchers (CITE HOWARD/CUNNINGHAM).
-However, our interpretation of the location of the thrusts is different from theirs.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-We did not investigate either of these faults in the field.
-However, the southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-There is not much evidence for the northern fault.
-However, the slip that created the uplifted topography on the northern side of the mountain probably occurs on some fault.
-To the east, the main trace of the Shargyn Fault enters the alluvial fan and disappears.
-Perhaps, this fault is uplifting the northern side of Little Dariv Mountain.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by igneous contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{offsetstreambedsshargyn.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-Although the granite intrusion is clearly a later event, the contact between the granite and the metamorphic is perfectly straight, supporting the interpretation of a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-There may be thrust faults that connect the Shargyn Fault with the Khantaishir thrusts, but we did not observe them in the field or on the satellite imagery.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-Though we did not observe the fault itself, the small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by authors (CITE THEM!).
-We examined the area.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south (CHECK this in field notebook) dipping high grade metamorphic rocks.
-Because of the topographic depression and depocenter, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-Further south along the cross-section, we encounter another fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-
-\section{Geophysical Data}
-	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
-
-	Bouguer gravity anomaly measurements (shown in Figure \ref{gravity}) indicate that the crust beneath the Shargyn Basin is denser than the surrounding area.
-Bouguer gravity anomalies are deviations from the geoid gravity field after correcting for the elevation of the measurement and the excess mass above the geoid.
-A deep crustal root implies lower density and thus, isostasy should cause large elevated regions to show a large negative bouguer anomaly, whereas topographically low regions should show a more positive bouguer anomaly.
-This can be seen in Figure \ref{gravity}, where the Altai mountain range and the Hangai mountain ranges show very strong negative Bouguer anomalies.
-The topographically low region south of the Altai and Gobi-Altai and the "Lakes" region, between the Altai and the Hangai both show significantly less negative anomalies, as expected for topographic lows.
-
-	Bouguer gravity anomaly measurements were calculated from the EGM2008 global gravity model \citep{EGM2008}, which uses a spherical harmonic expansion through order 2160.
-Compared to the surrounding mountain ranges, the Shargyn Basin shows a more negative bouguer anomaly.
-Furthermore, compared to the rest of the region, many of the mountain ranges immediately surrounding the Shargyn Basin show a surprisingly small negative bouguer anomaly.
-The negative bouguer anomaly beneath the Shargyn Basin can be explained by a denser crustal basement.
-The small negative bouguer anomaly in the surrounding mountain ranges may indicate lighter than average crustal densities in these regions.
-\citet{Bayasgalan2005a} calculates an average elastic thickness for Mongolia of ~6km.
-This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
-The density anomalies could be linked to the location of past and present faulting.
-Perhaps, the denser basement under the Shargyn Basin is more difficult to fracture and deform.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
-  \caption{Bouguer gravitational anomalies in western Mongolia.
-The Shargyn Basin is outlined in black.
-Compare the more negative anomalies under the Shargyn Basin to the more positive anomalies underneath the surrounding mountain ranges.
-This is the opposite of what we would expect from isostatic considerations.}
-  \label{gravity}
-\end{figure}
-
-	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
-The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
-The Shargyn Basin is at the location of a small positive magnetic anomaly.
-However, the spatial extent of the anomaly does not match very well with the shape of the Shargyn Basin.
-This is to be expected, because the minimum resolvable wavelength for a harmonic expansion of order 720 is ~56km.
-It is also likely that there is significant spatial error near the minimum wavelength.
-The Shargyn Basin is only 140km across -- less than 3 minimal wavelengths.
-Therefore, we are currently unable to decide whether there is a real magnetic anomaly beneath the Shargyn Basin.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 5in, height = 5.0in]{{magnetic.ps}.png}
-  \caption{Magnetic anomaly in western Mongolia. The Shargyn Basin is outlined in black. Note the much lower resolution than the gravity measurements.}
-  \label{magnetic}
-\end{figure}
-
-	Earthquake focal mechanisms provide useful information about active fault slip in the region.
-The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
-\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
-In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-	Seismic wave speeds underneath the basin could indicate different crustal properties.
-I investigated available P-wave and S-wave global teleseismic models, but no existing models have sufficient resolution to determine the wave speed anomaly beneath the Shargyn Basin.
-If there have been local seismological surveys performed, I am unaware of them.
-
-\section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
-
-	The first answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
-On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	As we've discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a different past tectonic environment in the Shargyn Basin.
-
-	Notably, we have explained a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting (cite cunningham 2005).
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.
-
-\section{Conclusion}
-	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
-Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
-Here, we studied the zone where the Altai fault system and the Gobi-Altai fault system intersect.
-Transpressional strike-slip deformation is dominant and results in discrete fault block mountain ranges with small basins in between.
-The Shargyn Basin is an anomalously large basin for the region and begs explaining.
-
-	We describe the basin as a combination of two active structures.
-A left-lateral compressional stepover on the Bogd Fault forms the Khantaishir range and the Shargyn Fault.
-A conjugate strike-slip intersection between the Bogd Fault and the Tonhil Fault forms a wedge of uplifted material in the southwest corner of the basin.
-Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
-
+\include{intro}
+\include{background}
+\include{structure}
+\include{geophysical}
+\include{basin}
+\include{conclusion}
 \bibliographystyle{plainnat}

### 2013-05-04 writing/paper/structure.tex
reformatted thesis, incorporating lots of oli's edits

\section{Structural Evidence}
The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover. The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin. The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts. However, some motion remains on the continuation of the Bogd Fault. The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge. Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range. Figure \ref{schematic} shows a schematic diagram of this structure.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
  \caption{Schematic structural fault map of the Shargyn Basin region. The extent is approximately the same as the extent in Figure \ref{regional}. Description of the structure is in the text. The shaded areas are basins. Unshaded areas are mountainous. The dotted lines represent the dominant foliation strike.}
  \label{schematic}
\end{figure}

	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts. The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault. On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust and the Shargyn Fault separates the low-lying basin from the adjacent mountain range while not accomodating the uplift.
\subsection{Remote Sensing}
Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers). I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery. Surface features indicating active motion were identified. Because the region is arid, many important features are clearly visible. For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature. Using such features from Landsat imagery was useful in identifying important features for further investigation in the field. Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.

In the remote sensing studies, I focused on identifying prominent faults using steep topography, sharp color changes (lithology changes), fault scarps, and offset features. I also mapped clearly distinguishable lithologic units using color change in the Landsat imagery. Where possible, I noted the strike of the foliation. Foliation is useful to map out the Paleozoic structural grain in the region. I also noted the location and distribution of alluvial fans. Locations that have alluvial fans forming on top of mountain ranges are important because that might indicate a time difference in the formation or slip on multiple faults.

\subsection{Field Mapping}
During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region. We spent part of the time focusing on the active faults. We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}). We mapped at a broad scale, but made careful note of features indicating recent brittle deformation. The level of fracturing and fault exposures were important in locating faults. Where no other evidence is available, a very steep topographic step is associated with fault. Mapping was performed on paper with the assistance of satellite imagery and a GPS device.

The results of this mapping are summarized in a fault map in Figure \ref{faultmap}. Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated. Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated. We performed new mapping surrounding the Shargyn Basin Eleven cross-sections were created. Six of these are shown in the text. The remaining cross-sections are in Appendix A. The following sections describe in detail the important structural results, while referencing these cross-sections.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts). The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections. The white boundaries separate different mapping the areas. The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed. The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}. The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
  \label{faultmap}
\end{figure}

\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 1.8in]{khantaishir_sat.png}
  \caption{Landsat imagery showing the Khantaishir Range and the mapped active faults.}
\end{figure}

The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT). The cross-section across these two faults is marked on the map as E-E'. Other minor faults are identifiable on the satellite imagery. We did not investigate the minor faults in the field. Both the UKT and LKT strike approximately NNW-SSE. The UKT is the northern fault. On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence with a thick conglomerate layer followed by shales, sandstones and carbonate reef deposits. At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present. No outcrop is found on the southwest side of the UKT. At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself. This zone is shown in Figure \ref{faultzones}A. Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$. Fracture dip was similar to the dip of the UKT itself. Rotated sigmoidal tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion. S-C fabric within the deformation zone also indicated thrust motion.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 6in]{faultzones.png}
  \caption{Photographs of exposed faults. Field workers are used as scale bars. A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'. B) An outcrop of the exposed thrust in cross-section G-G'}
  \ref{faultzones}
\end{figure}

The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT. South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium. Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and basalt flows, with interlayered sandy carbonate beds. Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium. This marks the fault trace for the LKT. The sedimentary sequence norteast of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?). The older alluvium further northwest also support a complex faulting history.

There is evidence that the motion on the LKT began after the motion on the UKT. In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons. Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed. The uplift was rapid enough that the drainage pattern was "locked in". Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase. Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.

\subsection{Shargyn Fault and Termination Thrust Splays}

The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault. Four of the transects focus on this fault. The fault strikes WSW-ENE. I have assumed that the fault is approximately vertical because the motion is strike-slip.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
  \caption{Cross sections through the Shargyn Fault and its termination splays.}
  \label{shargynfaultxsecs}
\end{figure}
\clearpage
The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}. The positive flower structure has been identified by previous researchers \citep{Howard2006}. Their interpretation is similar to the one discussed here. However, there are differences in the number and location of thrusts. To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin. Little Dariv Mountain is located on the boundary of these two basins. The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks. On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls. Separating the range is a thrust fault. The fault follows a valley that is clearly visible on satellite imagery. To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards. No other faults were identifiable in the core of Little Dariv Mountain.

On the cross-section, two other faults are indicated on the periphery of the mountain. The southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments. It may be related to a large anticline visible on the satellite imagery west of Little Dariv Mountain. The northern boundary fault is the inferred continuation of the main Shargyn Fault trace.  This continuation of the Shargyn Fault terminates as thrust motion on thrusts in the Dariv Basin. Because the Dariv Basin thrusts are the easternmost extent of the Sutai Range restraining bend \citep{Cunningham2003}, the thrusting involved in the Sutai Range restraining bend can be linked with the strike-slip motion on the Shargyn Fault.

In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}. To the north, the topography rises steeply towards a relatively flat plateau. The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by intrusive contacts. Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here. Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range. It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.

The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}. The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE). It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side. Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}). Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation. There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.  The fault itself passes through a topographic low and does not outcrop. On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation. Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade. The topography and lithology shifts both provide evidence for the presence of a fault at the marked location. The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming. The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 2.0in]{shargyn_fault_streams.png}
  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
  \label{shargynoffsetstreams}
\end{figure}

To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}. Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE). To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss. To the north of the fault, the topography is lower and transitions into alluvium. The northern unit is an slightly deformed alkali granite (show picture that I took?). The difference in metamorphic grade between the granite and the metamorphic unit indicates the granite was intruded after metamorphism. However, the contact between the granite and the metamorphic is perfectly straight, supporting the contact's interpretation as a fault. The topographic depression along the fault provides further evidence. To the east, the fault terminates somewhere near the city of Altai.

West of the transect, a thrust splays off the Shargyn Fault striking southwards. The thrust intersects the transect in the south between a red very poorly consolidated sandy unit and a steeply dipping shale. The northern side of the thrust is uplifted, while the basin is on the south side. Further south, elevation profiles suggest there may be much more thrust motion. Further south, this thrust connects with the Khantaishir Thrust system.

\subsection{Tonhil Fault and Intersection Wedge Thrusts}
Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
  \caption{Cross sections through the intersection wedge.}
  \label{intersectionwedgexsecs}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 1.8in]{intersection_wedge_sat.png}
  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin. Mapped active faults are shown by dashed lines Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.}
\end{figure}

At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}. After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south. However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault. There is a noticeable topographic rise bending away form the main Bogd Fault. The small depositional region and the topographic rise would be unlikely without recent motion. In addition, there was a higher local degree of damage to the bedrock. We observed a fracture set at (FILL IN S/D).

At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking. At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'. This has been previously identified on satellite imagery by \citet{Cunningham2010}. We examined the area in the field. Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise. North of the basin, we found a large granite intrusion. South of the basin, there are south dipping high grade metamorphic rocks. Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.

The western boundary of the intersection wedge is controlled by the Tonhil Fault. The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault. At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge. At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S. This bend appears to have created another transtensional basin. We did not examine the transtensional basin in the field. However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.

Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side. This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts. For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.

\subsection{Bogd Fault and Termination Thrust Splays}

Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault. These faults can be seen in the large scale fault map. Two transects were studied across these thrust splays.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
  \caption{Cross sections through the Shargyn Fault and its termination splays.}
  \label{bogdfaultxsecs}
\end{figure}

The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}. Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body. A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself. The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery. This is confirmed in the field, though minor folding interferes locally. This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit. The elevation along the basin margin decreases to the northwest, away from the Bogd Fault. This might indicate a continuous decrease in slip. Sediment accumulating behind the thrust points to the presence of another fault. Further south along the cross-section, we encounter this fault. The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence. Furthermore, there is another jump in topography at the fault. This fault was never observed. However, the satellite imagery indicates that it dominates the topographic expression west of our fault.

Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'. Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F'). This fault was visible in outcrop. The fault strikes approximately east and dips 15 degrees. The foot wall consists of low grade metamorphosed mudstones. The hanging wall consists of similar clastic metasedimentary rocks plus greenschist. Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge. Kinematic indicators suggest thrust sense of motion. A photograph is shown in Figure \ref{faultzones}B. The topography rises sharply directly south of the location of the fault. The basin margin elevation on this fault also decreases in the north suggesting a progressive decrease in slip.

### 2013-05-07 writing/paper/abstract.tex
writing.. writing.. writing..

@@
 The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-I examine the fault kinematics surrouding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
-The basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays.
+We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
+The studies were performed using satellite data and targeted field transects.
+The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays.
 The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong.

### 2013-05-07 writing/paper/background.tex
writing.. writing.. writing..

@@
 Important locations or faults are labeled.
+HTF = Hoh-Serh-Tsagaan-Salaa Fault, BF = Bolnay Fault, BGF = Bogd Fault.
 The study area and location of Figure \ref{regional} is outlined in black.
@@
 Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}.
+The geologic slip rate for a major transpressional fault, the Hoh-Serh-Tsagaan-Salaa Fault, is ~0.3 mm/yr of shortening and ~0.9 mm/yr of dextral slip.
+This consitutes ~20\% of the total strain accumulation in the Altai.
+\citep{Frankel2010}
 
@@
 The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System \citep{Cunningham2003a}.
+At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}.
 There are no slip estimates for the Bogd Fault System in the north.
@@
 	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
+	\includegraphics[width = 6in, height = 4in]{intro.png}
 	\caption{Shaded relief map showing notable features in the Shargyn Basin region.

### 2013-05-07 writing/paper/intro.tex
fixing up the maps

@@
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
-	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
-Important locations or faults are labeled.
-The study area and location of Figure \ref{regional} is outlined in black.
-The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important.
-Modified after \citet{Calais2003}}
-	\label{broad}
-\end{figure}
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = 6in, height = 4in]{regional.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region.
-Many of these faults are studied in this paper.
-GPS velocity magnitude for the Altai site is 4.4mm/yr.
-Faults are shown.
-See Figure \ref{faultmap} for the direction of slip.}
-	\label{regional}
-\end{figure}
-
 	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.

### 2013-05-07 writing/paper/structure.tex
writing.. writing.. writing..
fixing up the maps

@@
   \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts).
-The red lines indicate cross-sections and the adjacent letters provide an index for these cross-sections.
-The white boundaries separate different mapping the areas.
-The southern area is the mapping performed during this field season, with the smaller enclosed boxes being the areas which we directly observed.
-The northwestern mapping area of the Sutai Range was performed by \citet{Cunningham2003}.
-The northeastern mapping was performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.}
+  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts). }
   \label{faultmap}
@@
 Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwesternmost of the intersection wedge thrusts.
+This is interpreted as the furthest northern extent of the southwestern-most of the intersection wedge thrusts.
 For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
+
+ADD A SECTION ON THE MAIN TRANSECT THROUGH THE WEDGE?!
 

### 2013-05-08 writing/paper/abstract.tex
sending it to oli

@@
 We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
-The studies were performed using satellite data and targeted field transects.
-The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many terminal thrust splays.
+The studies were performed using satellite data and targeted field work transects.
+The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many strike-slip terminating thrust splays.
 The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong.

### 2013-05-08 writing/paper/background.tex
fixing up the maps
sending it to oli

@@
 	\centering
-	\includegraphics[width = 6.6in, height = 6in]{asia.png}
+	\includegraphics[width = 7in, height = 6in]{asia.png}
 	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
@@
 BHb, Bohaiwan basin; BSb, Banda Sea basin; CKg, Central Kamchatka graben; ECSb, East China Sea basin; FMg, Fushun-Meikehou graben; Fr, Fenwei rift; HYr, Hetao-Yinchuan rift; Jb, Java basins; MGb, Mergui basin; MLb, Malay basin; OT, Okinawa Trough; Pb, Pattani basin; Sb, Sumatra basins; SCSb, South China Sea basin; SCSmb, South China Sea marginal basin; SHdsz, Sakhalin-Hokkaido dextral shear zone; SLb, Shantar-Liziansky basin; SXr, Shanxi rift; Yb, Yunqing basin; YSb, Yellow Sea basin; YYg, Yilan-Yitong graben; Zb, Zengmu basin; Zu, Zenghe uplift" \citep{Schellart2005}.
-I have made additions to the figure: The location of Figure \ref{broad} is outlined in Central Asia.
-Gray arrows are added to indicate the direction of motion.
-The zone with spaced right-slip faults are surrounded by a dashed line.}
+The location of Figure \ref{broad} is outlined in Central Asia.
+}
 	\label{asia}
@@
 
-	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}(cite more).
+	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}\citep{Walker2007}\citep{Cunningham2005a}\citep{England1997}.
 Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models.
-The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citet{Tapponier1982}\citep{Peltzer1996}.
+The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citet{Tapponnier1982}\citep{Peltzer1996}.
 This hypthesis suggests that most motion has been accomodated by large strike-slip and thrust faults.
@@
 	Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponier1982}\citep{Peltzer1996}\citep{Houseman1993}.
+The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982}\citep{Peltzer1996}\citep{Houseman1993}.
 But, because of its orientation, active EASS rollback would exert stresses in a similar orientation to the Indo-Eurasian collision.
@@
 One explanation invokes bookshelf-like rotation along a long left-lateral shear zone creates the right-lateral faulting.
-\citep{Davy1998} \citep{Bayasgalan2005a}  This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
+\citep{Davy1988}\citep{Bayasgalan2005a}  This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
 Possibly, the Altai and Gobi-Altai are similar to an earlier phase in the evolution of the Tian Shan, given the progressive northward propagation of stresses as the India-Eurasian collision proceeds.
@@
 	\centering
-	\includegraphics[width = 6.5in, height = 4in]{broad.png}
+	\includegraphics[width = 7.0in, height = 4.09in]{broad.png}
 	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
@@
 HTF = Hoh-Serh-Tsagaan-Salaa Fault, BF = Bolnay Fault, BGF = Bogd Fault.
-The study area and location of Figure \ref{regional} is outlined in black.
+The study area and location of Figure \ref{intro} is outlined in black.
 The difference between right-lateral transpressional strike-slip in the Altai Range and left-lateral transpressional strike-slip in the Gobi-Altai is especially important.
@@
 The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the faults bounding the range to the southwest and west towards the Shargyn Basin.
-The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, is also poorly understood.
+However, the mapping from these studies has not seriously examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.
+The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, are also poorly understood.
 
@@
 	\centering
-	\includegraphics[width = 6in, height = 4in]{intro.png}
-	\caption{Shaded relief map showing notable features in the Shargyn Basin region.
-Many of these faults are studied in this paper.
-GPS velocity magnitude for the Altai site is 4.4mm/yr.
-Faults are shown.
-See Figure \ref{faultmap} for the direction of slip.}
-	\label{regional}
+	\includegraphics[width = 7in, height = 4.2in]{intro.png}
+	\caption{Landsat Imagery basemap showing important locations for this study.
+The large white box surrounding the Shargyn Basin is the new mapping from this study.
+The white area surrounding the Dariv Range indicates mapping performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.
+The area surrounding the Sutai Range was mapped by \citet{Cunningham2003}.
+The small white boxes around the Shargyn Basin show areas that we studied in the field.
+The red lines are labeled cross-sections that will be discussed in the Structural Evidence section.
+A GPS vector near the city of Altai is shown in the northeast.
+Earthquake focal mechanisms in the area are also shown.}
+	\label{intro}
 \end{figure}
@@
   \centering
-  \includegraphics[width = 4.6in, height = 4in]{basins.png}
+  \includegraphics[width = 4.5in, height = 5.67in]{basins.png}
   \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
@@
 \begin{itemize}
-	\item Foreland basins form in front a simple thrust system. Flexural subsidence may result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
+	\item Foreland basins form in front of a simple thrust system. Flexural subsidence can result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
 	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
@@
 	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Overview section.
+A summary of the results is in the Structural Interpretation section.
 Nearby regions have already been investigated.
@@
 The faults bounding the range exhibit a positive flower structure such that the faults on the east dip to the west and the faults on the west dip to the east\citep{Cunningham2003}.
-To the east, the Dariv Thrust uplifts the Dariv Mountains.
+
+	To the east of the Sutai Range, the Dariv Thrust uplifts the Dariv Mountains.
 \citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
@@
 
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different age.
+The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different ages.
 North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
@@
 Another phase of regional magmatism may have occured during the Permian \citep{Windley2007}.
-Small, post-orogenic sedimentary units are also present throughout the region.
+Small, post-orogenic sedimentary basins are also present throughout the region.
 
@@
   \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrock.png}
-  \caption{Schematic tectonostraigraphic map of Mongolia.
+  \includegraphics[width = 6.4in, height = 4.5in]{bedrockfigure.png}
+  \caption{Schematic tectonostratigraphic map of Mongolia.
 The area we studied is located in the outlined box.
@@
 The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-It is not shaded because it is not shown to be rigid like the Junggar Block.
+It is not shaded because it is probably not so mechanically rigid.
 Modified after \citet{Cunningham2005a}}
@@
 There is a close correspondence between the major faults and this structural fabric.
-The Main Mongolian Lineament could be a primary control on the location of the Bogd Fault.
 

### 2013-05-08 writing/paper/basin.tex
sending it to oli

@@
 \section{Formation of the Shargyn Basin}
-	It is interesting to ask why the Shargyn Basin exists.
+	With a kinematic description of the Shargyn Basin in hand, it is interesting to ask why the Shargyn Basin exists.
 The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
@@
 The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
-Other researchers have mapped the Main Mongolian Lineament as bending southwards in this area.
+\citet{Windley2007} has mapped the Main Mongolian Lineament as bending southwards in this area as well.
 On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
 
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes subparallel to the important fault in the region.
-The foliations could be explained by dragging along the fault.
+	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes parallel or close to parallel to the important fault in the region.
+The foliations fit well with a model where faults form along structural weaknesses.
+The foliations could also be explained by dragging along the fault.
 With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
 Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-But, in other locations, such dragging is not visible.
-The foliations also fit well with a model where faults form along structural weaknesses.
-Faults forming along weaknesses and dragging probably both contribute to the fault-parallel foliations.
+However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault and dragging is not seen.
+So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.
 
 	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a clearly positive density anomaly beneath the Shargyn Basin.
+Based on the bouguer gravity measurements described, there is a positive density anomaly beneath the Shargyn Basin.
 Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
@@
 The Tarim basin is much larger.
-The Tarim basin is thought to be a cratonal block, leading to its strength.
+The Tarim basin is an Archaen and Proterozoic cratonal block, leading to its strength\citep{Lu2008}.
 The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan.
+The northern border of the Tarim Basin is the Tian Shan Range.
 The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
@@
 
-	As has been discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-But, the Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins.
-Because of the complex Paleozoic tectonic environment, it is not implausible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-We cannot examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure, the Main Mongolian Lineament, in the area does not intersect the Shargyn Basin.
+	Because of the complex Paleozoic tectonic environment, it is possible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
+The Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins, does not intersect region Paleozoic structures, and exhibits a gravity anomaly.
+As discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
+We also do not have the data to examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure, the Main Mongolian Lineament, in the area does not intersect the Shargyn Basin.
 The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
@@
 
-	Notably, we have explained a large basin without any extension.
+	The three earthquake locations shown in \ref{eqlocs} that are in the Shargyn Basin could shed some doubt on this hypothesis concerning the basin's underlying basement.
+However, none of these earthquakes were particularly large.
+So, because of the low seismometer density in Central Asia, the error on the locations is large.
+If the earthquakes are, in fact, on active faults in the Shargyn Basin, because of the lack of topographic expression, the faults cannot have accomodated a particularly large amount of uplift.
+
+	Notably, the kinematics for the Shargyn Basin explain a large basin without any extension.
 Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.

### 2013-05-08 writing/paper/conclusion.tex
sending it to oli

@@
 	The India-Eurasia collision causes deformation through a broad swath of Asia.
-In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate.
+In central Asia, the Altai Range and Gobi-Altai Range accomodate 10-15\% of the total convergence rate \citep{Calais2003}.
 Hence, western Mongolia has some of the best examples in the world of intracontinental deformation patterns.
@@
 Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
+Thus, the structure Shargyn Basin region is a prime example of the features that are common in intraplate transpressional orogenies.
 
@@
 Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
+This description of the neotectonics of the Shargyn Basin emphasizes the importance of pre-existing crustal weakness and heterogeneities in controlling intraplate deformation.

### 2013-05-08 writing/paper/geophysical.tex
sending it to oli

@@
 	The lack of faulting in the Shargyn Basin is anomalous and could be due to unique basement geology.
-I derive constraints on the subsurface geology of the Shargyn Basin from geophysical measurements.
-I examined the available gravity, magnetic and seismic data for the region.
+I examined the available gravity, magnetic and seismic data for the region to derive constraints on the subsurface geology in the Shargyn Basin.
 
@@
 This elastic thickness would lead to almost complete isostatic adjustment for an area the size of the Shargyn Basin.
-Hence, the anomaly cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
+Hence, the anomaly probably cannot be attributed to isostatic adjustment from other nearby anomalies, like the Sutai Range.
 The density anomalies could be linked to the location of past and present faulting.
@@
   \centering
-  \includegraphics[width = 5in, height = 5.0in]{{gravity.ps}.png}
+  \includegraphics[width = 5in, height = 5.0in]{gravity.png}
   \caption{Bouguer gravitational anomalies in western Mongolia.
@@
 	Magnetic anomalies could also indicate a different composition of crust beneath the Shargyn Basin, however the signal is ambiguous in the Shargyn Basin.
-I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$.
+I used the NGDC-720 Crustal Magnetic Field Model, a elliptical harmonic expansion of the gravitational field to $n = 720$\citep{Maus2010}.
 The gravitational anomaly above the Shargyn Basin and surrounding areas is shown in Figure \ref{magnetic}.
@@
 	Earthquake focal mechanisms provide useful information about active fault slip in the region.
+The available calculated focal mechanisms for western Mongolia are shown in Figure \ref{broad}.
 The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 5in, height = 2.88in]{eq_locs.png}
+  \caption{Shaded relief map of the Shargyn Basin showing blue plus signs where earthquakes have occured between 1960 and 2013. Black lines are fault locations, like in Figure \ref{faultmap}.}
+  \label{eqlocs}
+\end{figure}	
+	Earthquake locations can also constrain the location of active deformation.
+For smaller earthquakes, focal mechanisms cannot be calculated because there is a lack of data.
+However, locations can still be determined, albeit with significantly decreased accuracy.
+The locations for all the earthquakes available from the IRIS Data Management System in the Shargyn Basin region are shown in Figure \ref{eqlocs}.
+Most of these earthquakes occur around the margins of the basin at locations near where we have mapped faults.
+Notably, there are three earthquakes shown in the eastern part of the Shargyn Basin.
+These earthquakes might indicate incipient faulting in the basin.
+
 	Seismic wave speeds underneath the basin could indicate different crustal properties.

### 2013-05-08 writing/paper/interpretation.tex
sending it to oli

--- 2013-05-07 writing/paper/structure.tex
@@
-\section{Structural Evidence}
+\section{Structural Interpretation}
 The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
@@
   \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
+  \caption{Schematic fault map of the Shargyn Basin region.
 The extent is approximately the same as the extent in Figure \ref{regional}.
@@
 Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
+The dotted lines represent the dominant foliation strike.
+}
   \label{schematic}
@@
 On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust and the Shargyn Fault separates the low-lying basin from the adjacent mountain range while not accomodating the uplift.
-\subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
-Surface features indicating active motion were identified.
-Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
-Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
-
-In the remote sensing studies, I focused on identifying prominent faults using steep topography, sharp color changes (lithology changes), fault scarps, and offset features.
-I also mapped clearly distinguishable lithologic units using color change in the Landsat imagery.
-Where possible, I noted the strike of the foliation.
-Foliation is useful to map out the Paleozoic structural grain in the region.
-I also noted the location and distribution of alluvial fans.
-Locations that have alluvial fans forming on top of mountain ranges are important because that might indicate a time difference in the formation or slip on multiple faults.
-
-\subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region.
-We spent part of the time focusing on the active faults.
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
-We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-The level of fracturing and fault exposures were important in locating faults.
-Where no other evidence is available, a very steep topographic step is associated with fault.
-Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
-
-The results of this mapping are summarized in a fault map in Figure \ref{faultmap}.
-Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-We performed new mapping surrounding the Shargyn Basin Eleven cross-sections were created.
-Six of these are shown in the text.
-The remaining cross-sections are in Appendix A.
-The following sections describe in detail the important structural results, while referencing these cross-sections.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts). }
-  \label{faultmap}
-\end{figure}
-
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir_sat.png}
-  \caption{Landsat imagery showing the Khantaishir Range and the mapped active faults.}
-\end{figure}
-
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
-Other minor faults are identifiable on the satellite imagery.
-We did not investigate the minor faults in the field.
-Both the UKT and LKT strike approximately NNW-SSE.
-The UKT is the northern fault.
-On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence with a thick conglomerate layer followed by shales, sandstones and carbonate reef deposits.
-At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present.
-No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
-This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
-Fracture dip was similar to the dip of the UKT itself.
-Rotated sigmoidal tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion.
-S-C fabric within the deformation zone also indicated thrust motion.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
-  \caption{Photographs of exposed faults.
-Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
-\end{figure}
-
-The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and basalt flows, with interlayered sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence norteast of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
-
-There is evidence that the motion on the LKT began after the motion on the UKT.
-In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
-Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed.
-The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
-
-\subsection{Shargyn Fault and Termination Thrust Splays}
-
-The northern margin of the basin is separated from the Dariv Range by the Shargyn Fault.
-Four of the transects focus on this fault.
-The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
-\end{figure}
-\clearpage
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
-The positive flower structure has been identified by previous researchers \citep{Howard2006}.
-Their interpretation is similar to the one discussed here.
-However, there are differences in the number and location of thrusts.
-To the north of the Shargyn Basin is the Dariv Basin, a much smaller, higher elevation basin.
-Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
-On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
-Separating the range is a thrust fault.
-The fault follows a valley that is clearly visible on satellite imagery.
-To the south of the fault is a steep 200 meter topographic rise, indicating that the fault dips southwards.
-No other faults were identifiable in the core of Little Dariv Mountain.
-
-On the cross-section, two other faults are indicated on the periphery of the mountain.
-The southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-It may be related to a large anticline visible on the satellite imagery west of Little Dariv Mountain.
-The northern boundary fault is the inferred continuation of the main Shargyn Fault trace.
-This continuation of the Shargyn Fault terminates as thrust motion on thrusts in the Dariv Basin.
-Because the Dariv Basin thrusts are the easternmost extent of the Sutai Range restraining bend \citep{Cunningham2003}, the thrusting involved in the Sutai Range restraining bend can be linked with the strike-slip motion on the Shargyn Fault.
-
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
-To the north, the topography rises steeply towards a relatively flat plateau.
-The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by intrusive contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
-Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
-It is more likely that the Dariv Thrust to the northeast is uplifting the range and the Shargyn Fault is simply accomodating the juxtaposition of the uplifted Dariv Range and the Shargyn Basin.
-
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 4in, height = 2.0in]{shargyn_fault_streams.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
-  \label{shargynoffsetstreams}
-\end{figure}
-
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
-To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
-To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
-The difference in metamorphic grade between the granite and the metamorphic unit indicates the granite was intruded after metamorphism.
-However, the contact between the granite and the metamorphic is perfectly straight, supporting the contact's interpretation as a fault.
-The topographic depression along the fault provides further evidence.
-To the east, the fault terminates somewhere near the city of Altai.
-
-West of the transect, a thrust splays off the Shargyn Fault striking southwards.
-The thrust intersects the transect in the south between a red very poorly consolidated sandy unit and a steeply dipping shale.
-The northern side of the thrust is uplifted, while the basin is on the south side.
-Further south, elevation profiles suggest there may be much more thrust motion.
-Further south, this thrust connects with the Khantaishir Thrust system.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{intersection_wedge_sat.png}
-  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
-Mapped active faults are shown by dashed lines Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.}
-\end{figure}
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-The small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by \citet{Cunningham2010}.
-We examined the area in the field.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south dipping high grade metamorphic rocks.
-Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwestern-most of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-ADD A SECTION ON THE MAIN TRANSECT THROUGH THE WEDGE?!
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-The elevation along the basin margin decreases to the northwest, away from the Bogd Fault.
-This might indicate a continuous decrease in slip.
-Sediment accumulating behind the thrust points to the presence of another fault.
-Further south along the cross-section, we encounter this fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-The basin margin elevation on this fault also decreases in the north suggesting a progressive decrease in slip.

### 2013-05-08 writing/paper/intro.tex
sending it to oli

@@
 Furthermore, where multiple fault systems intersect, complex kinematics can occur.
-Understanding how and why these deformation patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
+Understanding how and why these kinematic patterns develop is important for understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
 

### 2013-05-08 writing/paper/structure.tex
sending it to oli

@@
 \section{Structural Evidence}
-The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
-The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
-The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
-However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
-Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
-Figure \ref{schematic} shows a schematic diagram of this structure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
-  \caption{Schematic structural fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
-Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
-The dotted lines represent the dominant foliation strike.}
-  \label{schematic}
-\end{figure}
-
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust and the Shargyn Fault separates the low-lying basin from the adjacent mountain range while not accomodating the uplift.
 \subsection{Remote Sensing}
-Previous studies have examined Mongolian active tectonics from a remote sensing perspective (cite some of the tapponnier papers and the cunningham papers and walker papers).
+Previous studies have examined Mongolian active tectonics from a remote sensing perspective \citep{Cunningham2005a}\citep{Cunningham2010}\citep{Tapponnier1979}\citep{Walker2007}.
 I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
@@
 Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
-Also, using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
+Using contiguous features on the satellite imagery, field mapping results can be extended to nearby, unobserved areas.
 
 In the remote sensing studies, I focused on identifying prominent faults using steep topography, sharp color changes (lithology changes), fault scarps, and offset features.
-I also mapped clearly distinguishable lithologic units using color change in the Landsat imagery.
+I also mapped clearly distinguishable lithologic units using color, weathering, and foliation changes visible in the Landsat imagery.
 Where possible, I noted the strike of the foliation.
@@
 I also noted the location and distribution of alluvial fans.
-Locations that have alluvial fans forming on top of mountain ranges are important because that might indicate a time difference in the formation or slip on multiple faults.
+Locations where alluvial fans are forming above active thrusts are important because that might indicate a temporal or slip difference between multiple adjacent faults.
 
@@
 We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
-The level of fracturing and fault exposures were important in locating faults.
-Where no other evidence is available, a very steep topographic step is associated with fault.
+Topography, fracturing, fault exposures and shifts in foliations or lithology were all important in locating faults.
 Mapping was performed on paper with the assistance of satellite imagery and a GPS device.
+Later, the mapping was transferred to digital form.
 
@@
 Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
-We performed new mapping surrounding the Shargyn Basin Eleven cross-sections were created.
+We performed new mapping surrounding the Shargyn Basin.
+Eleven cross-sections were created.
 Six of these are shown in the text.
@@
   \centering
-  \includegraphics[width = 7in, height = 5.0in]{faultmap.png}
-  \caption{The black lines are faults with the indicated slip direction and dip direction (for thrusts). }
+  \includegraphics[width = 7in, height = 5.0in]{fault_map.png}
+  \caption{A map of the Shargyn Basin showing the studied faults and important locations. Slip is indicated by thrust teeth or lateral slip arrows.}
   \label{faultmap}
@@
 
-\subsection{Khantaishir Thrusts -- Northwards Slip Transfer}
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir.png}
-  \caption{Northeast-Southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts.}
-\end{figure}
+\subsection{Khantaishir Thrusts}
 
@@
   \centering
-  \includegraphics[width = 6in, height = 1.8in]{khantaishir_sat.png}
-  \caption{Landsat imagery showing the Khantaishir Range and the mapped active faults.}
+  \includegraphics[width = 7in, height = 9.42in]{khantaishir_joint.png}
+  \caption{A northeast-southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts paired with a Landsat image.
+Important locations and lithology are noted on the satellite image.}
 \end{figure}
 
-The eastern edge of the basin is uplifted with motion concentrated on two fault which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
-The cross-section across these two faults is marked on the map as E-E'.
+The eastern edge of the basin is uplifted with motion concentrated on two faults which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT).
+The cross-section across these two faults is marked on the map as A-A'.
 Other minor faults are identifiable on the satellite imagery.
@@
 No outcrop is found on the southwest side of the UKT.
-At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that could be interpreted as the UKT itself.
+At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that we interpret as the UKT itself.
 This zone is shown in Figure \ref{faultzones}A.
@@
   \centering
-  \includegraphics[width = 4in, height = 6in]{faultzones.png}
+  \includegraphics[width = 4in, height = 6in]{faultscollage.png}
   \caption{Photographs of exposed faults.
@@
 B) An outcrop of the exposed thrust in cross-section G-G'}
-  \ref{faultzones}
+  \label{faultzones}
 \end{figure}
@@
 This marks the fault trace for the LKT.
-The sedimentary sequence norteast of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area (CITE the russian map which called it devonian?).
-The older alluvium further northwest also support a complex faulting history.
+The sedimentary sequence northeast of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area.
 
-There is evidence that the motion on the LKT began after the motion on the UKT.
+Motion on the LKT may have begun after the motion on the UKT.
 In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons.
@@
 The uplift was rapid enough that the drainage pattern was "locked in".
-Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the pre-uplift phase.
-Southwestwards drainage in the pre-uplift phase is compatible with already ongoing uplift on the UKT.
+Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the time-period before the LKT was active.
+The deeply drainages between the LKT and the UKT also indicate that the LKT is a later uplift.
+A major drainage cuts through the uplift adjacent to the LKT.
+Such a drainage would be unlikely to form unless it was present before the activation of the LKT, providing further evidence that uplift began on the UKT before beginning on the LKT.
+
+\subsection{Bogd Fault and Termination Thrust Splays}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 7in, height = 6.83in]{haliun_joint.png}
+  \caption{A cross-section through the Haliun Thrust and Butara Thrust paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
+  \label{haliun_joint}
+\end{figure}
+
+Along the southern margin of the basin, we performed two transects across two thrusts splayed off the Bogd left-lateral strike slip fault.
+We will refer to the northern thrust as the Haliun Thrust and the southern thrust as the Butara Thrust.
+Above (south) of the Haliun Thrust, there is a complex sequence of metasedimentary and metavolcanics rocks that are intruded by a large gabbro body.
+There might be a fault-adjacent anticline.
+I do not know whether such an anticline is actively forming or a remnant from previous orogenies.
+The foliation in the fault-adjacent units is parallel to the Haliun Thrust and dips to the south.
+Along the basin margin, elevations decrease to the west.
+The decrease in elevation indicates the uplift tapers off further from the Bogd strike-slip fault.
+
+To the south of the Haliun Thrust, the Butara Thrust uplifts 3000m high peaks in the center of the Bogd Range.
+To the north of the Butara Thrust, we observed the same metasedimentary sequence as is adjacent to the Haliun Thrust.
+To the south of the Butara Thrust, there are sandstones and basaltic units.
+Just to the west of the cross-section, alluvium is accumulating south of the Haliun Thrust, indicating that more of the uplift is occuring on the Butara Thrust.
+To the west of cross-section, the high topography is almost entirely south of the Butara Thrust.
+
+The Butara Thrust is exposed 25km west of cross-section B-B'.
+Data from a transect performed at this location is shown in cross-section X1-X1' in Appendix A.
+Based on the outcrop, the fault strikes E-W and dips 15$\circ$.
+The foot wall consists of low grade metamorphosed mudstones and sandstone.
+The hanging wall consists of greenschist.
+Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
+Kinematic indicators suggest thrust sense of motion.
+A photograph is shown in Figure \ref{faultzones}B.
+The topography rises sharply directly south of the location of the fault.
+
+Further west, there is a third thrust splay off the Bogd Fault that can be seen in Figure \ref{faultmap}.
+But, we did not visit this fault in the field map.
+The satellite imagery suggests that it is similar in nature to the Haliun Thrust and the Butara Thrust.
+
+\subsection{Tonhil Fault and Intersection Wedge Thrusts}
+On the south-western margin of the Shargyn Basin, the wedge between the Tonhil Fault and Bogd Fault exhibits some of the most extensive thrusting.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 7in, height = 8.88in]{intersection_wedge_joint.png}
+  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
+Mapped active faults are shown by solid red lines.
+Slip sense is shown with thrust teeth or lateral slip arrows.
+Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.
+Cross sections C-C' and D-D' through the transtensional basin and the intersection wedge are shown.
+Cross-sections X2-X2' and X3-X3' are shown in Appendix A.}
+  \label{intersection_wedge_joint}
+\end{figure}
+
+At the eastern corner of the intersection wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
+At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section C-C'.
+This has been previously identified on satellite imagery by \citet{Cunningham2010}.
+We examined the area in the field.
+Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
+North of the basin, we found a large granite intrusion.
+South of the basin, there are south dipping high grade metamorphic rocks.
+Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
+
+At cross-section D-D' (shown in Figure \ref{intersection_wedge_joint}, we performed a transect through the series of thrusts separating the basin from the intersection wedge.
+At the site of the transect, there are three thrusts.
+The lowest thrust marks the basin margin.
+To the southwest of this thrust, there is a large hornblende-rich diorite intrusion with minor intrusions of granite included.
+Close to the middle thrust, there is sedimentary sequence that includes an actively mined coal bed.
+The middle thrust causes a noticeable topographic jump.
+On the southwest side of the middle thrust, there are intrusions of gabbro and diorite.
+The southwesternmost thrust separates these intrusive rocks from a schist and carbonate unit.
+To the southwest, the topography levels out onto a broad plateau over the whole intersection wedge area.
+Based on the topography, each of these faults accomodates some portion of the intersection wedge uplift.
+However, digital elevation profiles separating the basin and the intersection wedge point to along-strike variation in how much uplift is accomodated on each of the faults.
+
+The western boundary of the intersection wedge is controlled by the Tonhil Fault.
+The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
+At the cross-section X3-X3' (shown in Appendix A), we examine the fault where it interacts with the northwestern corner of the intersection wedge.
+At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
+This bend appears to have created another transtensional basin.
+We did not examine the transtensional basin in the field.
+It is intriguing that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section C-C'.
+
+Further east, along cross-section X3-X3', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
+This is interpreted as the furthest northern extent of the southwestern-most of the intersection wedge thrusts.
+For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
+
+At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section X2-X2' in Appendix A.
+After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
+However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
+There is a noticeable topographic rise bending away form the main Bogd Fault.
+The small depositional region and the topographic rise would be unlikely without recent motion.
+In addition, there was a higher local degree of damage to the bedrock.
 
@@
 The fault strikes WSW-ENE.
-I have assumed that the fault is approximately vertical because the motion is strike-slip.
+In the cross-sections, I have assumed that the Shargyn Fault is approximately vertical because of its lateral slip and the lack of other constraints.
 
@@
   \centering
-  \includegraphics[width = 6in, height = 9.0in]{shargynfault.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{shargynfaultxsecs}
+  \includegraphics[width = 7in, height = 7.5in]{lil_dariv_joint.png}
+  \caption{A cross-section through the Little Dariv Mountain area paired with a Landsat image of the surrounding area. Important locations and lithology are noted on the satellite image.}
+  \label{lil_dariv_joint}
 \end{figure}
 \clearpage
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section A-A' in Figure \ref{shargynfaultxsecs}.
+The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section E-E' in Figure \ref{lil_dariv_joint}.
 The positive flower structure has been identified by previous researchers \citep{Howard2006}.
@@
 Little Dariv Mountain is located on the boundary of these two basins.
-The majority of the range is composed of greenschist-facies sedimentary and volcanic rocks.
+The majority of the mountain is composed of greenschist-facies sedimentary and volcanic rocks.
 On the northern side of the range, there is a 100 meter thick band of folded carbonates and marls.
@@
 
-In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked B-B' in Figure \ref{shargynfaultxsecs}.
+In the central south of the Dariv Range, we examined a simple section of the Shargyn Fault, marked X4-X4' in Figure \ref{intro} and shown in Appendix A.
 To the north, the topography rises steeply towards a relatively flat plateau.
 The bedrock is composed of serpentinite, biotite-gabbro and gabbro-diorite, separated by intrusive contacts.
-Despite the steep topographic rise at the fault, it is not likely there is much thrusting on the Shargyn Fault here.
+Despite the steep topographic rise at the fault, it is unlikely that there is any dip-slip on the Shargyn Fault here.
 Thrusting on the Shargyn Fault at this location would conflict with the lack of topographic expression along the fault east of the Dariv Range.
@@
 
-The simplest cross-section of the Shargyn Fault is in the northeastern part of the basin, marked on the map as C-C' and shown in Figure \ref{shargynfaultxsecs}.
-The Shargyn fault is visible in satellite imagery of the cross-section area as a shift in topographic expression (REFERENCE IMAGE).
-It also marks a shift from a black unit on the south side to a dark and light brown unit on the north side.
-Slightly to the west of the cross-section, in the basin, there are clearly visible sinistral stream offsets (Figure \ref{shargynoffsetstreams}).
-Field observations indicate that the southern unit is primarily a quart-rich greenschist with a northwards dipping foliation.
-There are also minor components of metamorphosed conglomerate with preserved clasts and carbonate-rich rocks, especially closer to the fault.
-The fault itself passes through a topographic low and does not outcrop.
-On the north side of the fault, we see a higher-grade gneissic unit with steeper, south dipping foliation.
-Further north, we found an outcrop of garnet schist, confirming the shift in metamorphic grade.
-The topography and lithology shifts both provide evidence for the presence of a fault at the marked location.
-The topography (maybe show a picture?) and the offset stream beds (definitely show an image of the offset stream beds) located to the west imply that the Shargyn Fault is actively deforming.
-The slight southward dip of the fault was not observed but is inferred based on the elevated topography on the southern side of the fault.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 7in, height = 6.85in]{neshargyn_joint.png}
+  \caption{A cross-section through the Shargyn Fault in the northeast corner of the Shargyn Basin paired with a Landsat image of the surrounding area.
+Important locations and lithology are noted on the satellite image.}
+  \label{neshargyn_joint}
+\end{figure}
+
+In the northeastern part of the basin, the Shargyn Fault and a small adjacent thrust are shown in cross-section F-F' (Figure \ref{neshargyn_joint}).
+The small thrust uplifts the area south of the Shargyn Fault.
+The thrust is visible as a small, potentially recent \citep{Bayasgalan2005a}, scarp in the alluvium.
+The Shargyn Fault marks a major lithologic change in this area.
+To the south of the Shargyn Fault, we find quartz-rich greenschist, metabasalt and carbonate units with foliations dipping to the north.
+On the north side of the fault, there are amphibolite facies gneiss and schist units with south-dipping foliations.
+There are also some large granitoid intrusions.
+Outcrops of garnet schist confirm the higher metamorphic grade.
+To the west, offset stream beds shown in Figure \ref{shargynoffsetstreams} provide one of the clearest indications of the active slip sense.
 
@@
 
-To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked D-D' in Figure \ref{shargynfaultxsecs}.
-Here, the fault is still noticeable in satellite imagery (REFERENCE IMAGE).
+To the east, the strike of the Shargyn fault bends slightly southwards, at the cross-section marked X5-X5', shown in Appendix A.
+Here, the fault is still visible in satellite imagery as a eroded valley.
 To the south of the fault, we see steep rolling hills with highly deformed and metamorphosed amphibolite and gneiss.
 To the north of the fault, the topography is lower and transitions into alluvium.
-The northern unit is an slightly deformed alkali granite (show picture that I took?).
+The northern unit is an slightly deformed alkali granite.
 The difference in metamorphic grade between the granite and the metamorphic unit indicates the granite was intruded after metamorphism.
@@
 The northern side of the thrust is uplifted, while the basin is on the south side.
-Further south, elevation profiles suggest there may be much more thrust motion.
-Further south, this thrust connects with the Khantaishir Thrust system.
-
-\subsection{Tonhil Fault and Intersection Wedge Thrusts}
-Further west, the wedge between the Tonhil and Gobi-Altai Faults exhibits some of the most extensive thrusting.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 6.0in]{intersectionwedgexsecs.png}
-  \caption{Cross sections through the intersection wedge.}
-  \label{intersectionwedgexsecs}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 1.8in]{intersection_wedge_sat.png}
-  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
-Mapped active faults are shown by dashed lines Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.}
-\end{figure}
-
-At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section J-J' in Figure \ref{intersectionwedgexsecs}.
-After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south.
-However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault.
-There is a noticeable topographic rise bending away form the main Bogd Fault.
-The small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
-We observed a fracture set at (FILL IN S/D).
-
-At the eastern corner of the wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section H-H'.
-This has been previously identified on satellite imagery by \citet{Cunningham2010}.
-We examined the area in the field.
-Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise.
-North of the basin, we found a large granite intrusion.
-South of the basin, there are south dipping high grade metamorphic rocks.
-Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.
-
-The western boundary of the intersection wedge is controlled by the Tonhil Fault.
-The Tonhil Fault is a major NNW-SSE striking right-lateral strike slip fault.
-At the cross-section K-K', we examine the fault where it interacts with the northwestern corner of the intersection wedge.
-At the latitude of the cross-section, the Tonhil fault bends so that it strikes nearly N-S.
-This bend appears to have created another transtensional basin.
-We did not examine the transtensional basin in the field.
-However, it is interesting that it lies symmetrically across the intersection wedge from the transtensional basin on the Bogd Fault examined in cross-section H-H'.
-
-Further east, along cross-section K-K', we observed a boundary between greenschist grade metasedimentary rocks on the east side and a gneissic unit on the west side.
-This is interpreted as the furthest northern extent of the southwestern-most of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
-
-ADD A SECTION ON THE MAIN TRANSECT THROUGH THE WEDGE?!
-
-\subsection{Bogd Fault and Termination Thrust Splays}
-
-Along the southern margin of the basin, we encounter a repeated motif where thrusts splay off the main Gobi-Altai left-lateral strike slip fault.
-These faults can be seen in the large scale fault map.
-Two transects were studied across these thrust splays.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = 6in, height = 3.5in]{bogdfaultxsecs.png}
-  \caption{Cross sections through the Shargyn Fault and its termination splays.}
-  \label{bogdfaultxsecs}
-\end{figure}
-
-The easternmost of these thrusts can be seen in cross-section F-F' in Figure \ref{bogdfaultxsecs}.
-Directly above (south) the thrust, there is a complex sequence of metasedimentary and metavolcanic that is intruded by a large gabbro body.
-A bit to the west of the cross-section line, the topography lowers and outcrop is only visible within a short band near the fault itself.
-The average strike of the foliation in the fault-adjacent metasedimentary unit is visibly parallel to the fault on satellite imagery.
-This is confirmed in the field, though minor folding interferes locally.
-This suggests that the fault follows a pre-existing weakness in the meta-sedimentary unit.
-The elevation along the basin margin decreases to the northwest, away from the Bogd Fault.
-This might indicate a continuous decrease in slip.
-Sediment accumulating behind the thrust points to the presence of another fault.
-Further south along the cross-section, we encounter this fault.
-The lithology changes sharply from a sandstone and basalt sequence to a greenschist facies metasedimentary sequence.
-Furthermore, there is another jump in topography at the fault.
-This fault was never observed.
-However, the satellite imagery indicates that it dominates the topographic expression west of our fault.
-
-Twenty five kilometers to the west, we identified and visisted another of these thrust splays at cross-section G-G'.
-Based on continuous uplift, it appears to be the same fault as the one observed in the southern portion of the Haliun cross-section (F-F').
-This fault was visible in outcrop.
-The fault strikes approximately east and dips 15 degrees.
-The foot wall consists of low grade metamorphosed mudstones.
-The hanging wall consists of similar clastic metasedimentary rocks plus greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
-Kinematic indicators suggest thrust sense of motion.
-A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
-The basin margin elevation on this fault also decreases in the north suggesting a progressive decrease in slip.
+Elevation profiles suggest there the fault accomodates increasing amounts of uplift further south.
+At its southern end, this thrust connects with the Upper Khantaishir Thrust.

### 2013-05-08 writing/paper/xsecappendix.tex
sending it to oli

\section{Appendix A: More cross sections.}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 7.0in]{x1x2x3.png}
  \caption{Cross section X1 across the Butara Thrust on the southern margin of the Shargyn Basin, cross section X2 across the final terminal thrust of the Bogd Fault, and cross section X3 examining the Tonhil Fault and northwestern corner of the intersection wedge.}
  \label{lil_dariv_joint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 5.25in]{x4x5.png}
  \caption{Cross section X4 through the Shargyn Fault at the southern margin of the Dariv Range and cross section X5 at the eastern margin of the Shargyn Basin.}
  \label{lil_dariv_joint}
\end{figure}

### 2013-05-08 writing/paper/grabimages.py

@@ -1,37 +1,0 @@
-#!/usr/bin/python
-import os
-import subprocess
-def call_command(command):
-    process = subprocess.Popen(command.split(' '),
-                               stdout=subprocess.PIPE,
-                               stderr=subprocess.PIPE)
-    return process.communicate()
-
-call_command('../shargynxsecs/composites/shargynfault/make_composite.py')
-call_command('../shargynxsecs/composites/bogdfault/make_composite.py')
-call_command('../shargynxsecs/composites/intersectionwedge/make_composite.py')
-call_command('../shargynxsecs/composites/khantaishir/make_composite.py')
-
-
-files = dict()
-files['./shargynfault.png'] = '../shargynxsecs/composites/shargynfault/final.png'
-files['./bogdfaultxsecs.png'] = '../shargynxsecs/composites/bogdfault/final.png'
-files['./intersectionwedgexsecs.png'] = '../shargynxsecs/composites/intersectionwedge/final.png'
-files['./khantaishir.png'] = '../shargynxsecs/composites/khantaishir/final.png'
-files['./gravity.ps.png'] = '../gravity/gravity.ps.png'
-files['./magnetic.ps.png'] = '../magnetic/magnetic.ps.png'
-files['./broad.png'] = '../broadfiguremine/broad.png'
-files['./regional.png'] = '../regionalfigure/regional.png'
-files['./basins.png'] = '../basintypesfigure/basins.png'
-files['./asia.png'] = '../asiafigure/asia.png'
-files['./bedrock.png'] = '../bedrockfigure/bedrockfigure.png'
-files['./grain.png'] = '../grainfigure/grain.png'
-files['./faultzones.png'] = '../faultscollage/faultscollage.png'
-files['./schematic.png'] = '../schematicfaults/schematic.png'
-files['./faultmap.png'] = '../maps/markedupfaultmap.png'
-
-for key,value in files.iteritems():
-	cmd = 'cp ' + value + ' ' + key
-	print cmd
-	call_command(cmd)
-

### 2013-05-08 writing/paper/abstract.tex
removed sections on gravity

@@
 We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
-The studies were performed using satellite data and targeted field work transects.
+The studies were performed using satellite data and targeted field transects.
 The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many strike-slip terminating thrust splays.
-The fault-kinematics, pre-existing structural weaknesses and a significant gravity anomaly all suggest that the basin basement is mechanically strong.
+Furthermore, local foliation is almost always fault parallel, pointing to the importance of pre-existing structural weaknesses in the development of active faults.
 This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.

### 2013-05-08 writing/paper/conclusion.tex
removed sections on gravity

--- 2013-05-04 writing/paper/conclusion.tex
@@
 Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-
-	A density anomaly was identified underneath the Shargyn Basin using bouguer gravity anomaly data.
-Some difference in crustal basement may explain the density anomaly.
-Such crustal basement may be mechanically strong and difficult to fault.
-So, this unique block of crust may be controlling the deformation in the area.
-The lack of cross-cutting inactive faults indicates that most of the active deformation is occuring on ancient multiply-reactivated faults.
-Combined with the fact that the Paleozoic and Mesozoic structures in the area steer around the basin, we have further support for a unique crustal block.
+Thus, the structure Shargyn Basin region is a prime example of the features that are common in intraplate transpressional orogenies and the part that pre-existing structural weaknesses can play in defining their locations.

### 2013-05-08 writing/paper/discussion.tex
removed sections on gravity

--- 2013-05-08 writing/paper/basin.tex
@@
-\section{Formation of the Shargyn Basin}
-	With a kinematic description of the Shargyn Basin in hand, it is interesting to ask why the Shargyn Basin exists.
-The Upper and Lower Khantaishir Thrusts translate a portion of the strike-slip motion north into the Shargyn Fault.
-Westward (left-lateral) advection of material along the Bogd Fault is transformed into uplift by the Khantaishir thrusts.
-However, further north, this westward motion continues.
-The differential motion between this westward motion and the comparatively stationary Shargyn Basin is accomodated on the Shargyn Fault and the Dariv Thrust.
-But, why does this stepover form at all?
-Moreover, why did it form in its current position?
-What does the answer tells us about the tectonic processes at work in the area?
-These are all fundamentally the same question: what causes a given fault to form in a specific location?
-This is a deep and difficult question to answer in the general case.
-However, for the Shargyn Basin, there is a partial answer.
+\subsection{Interpretation and Discussion}
+	The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
+The Bogd Fault to the south, the Tonhil Fault to the west, the Shargyn Fault to the north and the Khantaishir Thrusts to the east are the primary faults governing the structure of the Shargyn Basin.
+The Bogd Fault and the Shargyn Fault are linked in a compressional stepover by the Khantaishir Thrusts.
+However, some motion remains on the continuation of the Bogd Fault.
+The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
+Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
+Figure \ref{schematic} shows a schematic diagram of this structure.
 
-	The first answer to this questions involves pre-existing weaknesses in the crust.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
+  \caption{Schematic fault map of the Shargyn Basin region.
+The extent is approximately the same as the extent in Figure \ref{regional}.
+Description of the structure is in the text.
+The shaded areas are basins.
+Unshaded areas are mountainous.
+The dotted lines represent the dominant foliation strike.
+}
+  \label{schematic}
+\end{figure}
+
+	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
+The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
+On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust and the Shargyn Fault separates the low-lying basin from the adjacent mountain range while not accomodating the uplift.
+
+	Notably, the kinematics for the Shargyn Basin explain a large basin without any extension.
+Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
+Previous researchers have stated that the region has no normal faulting or transtensional faulting \citep{Cunningham2005a}.
+The low elevations do not require extension.
+The "Valley of the Lakes" to the north is at similar elevations.
+The small transtensional basins near the intersection wedge were the only extensional features we found.
+However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
+Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
+These features are also much smaller than the Shargyn Basin.
+
+	With a kinematic description of the Shargyn Basin in hand, it is interesting to ask why such structures formed in their current locations.
+One answer to this questions involves pre-existing weaknesses in the crust.
 Near the stepover, the Bogd Fault bends southward 20-30 degrees.
@@
 So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.
-
-	The second answer to the question of why the Shargyn Basin formed involves the lithology of the basin basement.
-Based on the bouguer gravity measurements described, there is a positive density anomaly beneath the Shargyn Basin.
-Because identifying the exact source of a gravity anomaly is mathematically ill-posed, it is impossible to determine precisely what causes this positive density anomaly.
-But, it does indicate that something is different.
-The positive density anomaly combined with the lack of faults cutting the basin suggests a mechanically strong crustal block.
-The Sichuan and Tarim basins are other, larger, more well-known, examples of mechanically strong crustal blocks.
-
-	The Sichuan basin is underlain by a fast P-wave velocity zone down to 250 km \citep{Royden2008}.
-A very steep topographic rise surrounds the north and west margins of the topographically low basin.
-The Sichuan basin is up to 10km deep and ~300km E-W by 200km N-S in extent \citep{Burchfiel2008}.
-The Tarim basin is much larger.
-The Tarim basin is an Archaen and Proterozoic cratonal block, leading to its strength\citep{Lu2008}.
-The weak Tibetan Plateau is separated from the Tarim Basin by the Altyn Tagh Fault (see Figure \ref{asia}).
-The northern border of the Tarim Basin is the Tian Shan Range.
-The Tian Shan is especially mechanically weak, leading to the concentration of deformation.
-Both the Sichuan and Tarim basins show very little Mesozoic and Cenozoic deformation compared to their surroundings, indicating that their strength is not a recent development \citep{Neil1997}.
-The lack of active faults or topography cutting the basins is an obvious indication of their strength.
-The Junggar, 350km to the southwest of the Shargyn Basin, is also interpreted as a mechanically strong basin, perhaps due to a trapped piece of oceanic crust \citep{Carroll1990a}.
-
-	Because of the complex Paleozoic tectonic environment, it is possible for a large distinct crustal block to have accreted at the site of the Shargyn Basin.
-The Shargyn Basin is surrounded by uplifts, similar to the Tarim or Sichuan Basins, does not intersect region Paleozoic structures, and exhibits a gravity anomaly.
-As discussed, seismic velocity anomalies are too low resolution to determine anything about the Shargyn Basin.
-We also do not have the data to examine the Shargyn Basin for Paleozoic or Mesozoic deformation, but the largest Paleozoic structure, the Main Mongolian Lineament, in the area does not intersect the Shargyn Basin.
-The surface trace of the Main Mongolian Lineament follows the southern boundary of the Shargyn Basin from south of the Khantaishir Range west to the Tonhil Fault.
-The exact reason for this bend in the Main Mongolian Lineament is unknown, but it could indicate a regionally distinct past tectonic environment in the Shargyn Basin.
-
-	The three earthquake locations shown in \ref{eqlocs} that are in the Shargyn Basin could shed some doubt on this hypothesis concerning the basin's underlying basement.
-However, none of these earthquakes were particularly large.
-So, because of the low seismometer density in Central Asia, the error on the locations is large.
-If the earthquakes are, in fact, on active faults in the Shargyn Basin, because of the lack of topographic expression, the faults cannot have accomodated a particularly large amount of uplift.
-
-	Notably, the kinematics for the Shargyn Basin explain a large basin without any extension.
-Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
-Previous researchers have stated that the region has no normal faulting or transtensional faulting \citep{Cunningham2005a}.
-The low elevations do not require extension.
-The "Valley of the Lakes" to the north is at similar elevations.
-The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
-Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
-These features are also much smaller than the Shargyn Basin.
-In addition to these arguments, the low elevations could be due to the positive density anomaly.
-Denser crust would naturally isostatically adjust to be topographically low.

### 2013-05-08 writing/paper/structure.tex
removed sections on gravity

@@
 At its southern end, this thrust connects with the Upper Khantaishir Thrust.
+
+\subsection{Earthquake Mechanisms and Locations}
+	Earthquake focal mechanisms provide useful information about active fault slip in the region.
+The available calculated focal mechanisms for western Mongolia are shown in Figure \ref{broad}.
+The World Stress Map \citep{WorldStressMap2008} uses focal mechanisms as the exclusive constraint on the primary stress axis in western Mongolia due to the lack of other data.
+\citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
+Only a couple mechanisms are located in the Shargyn Basin.
+On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
+In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
+Both possibilities are reasonable given the location of the earthquake.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = 5in, height = 2.88in]{eq_locs.png}
+  \caption{Shaded relief map of the Shargyn Basin showing blue plus signs where earthquakes have occured between 1960 and 2013.
+Black lines are fault locations, like in Figure \ref{faultmap}.
+Note that many of the earthquakes occur on, or near, mapped faults.
+However, three of the earthquakes occured within the basin.}
+  \label{eqlocs}
+\end{figure}
+
+	Earthquake locations can also constrain the location of active deformation.
+For smaller earthquakes, focal mechanisms cannot be calculated because there is a lack of data.
+However, locations can still be determined, albeit with significantly decreased accuracy.
+The locations for all the earthquakes available from the IRIS Data Management System in the Shargyn Basin region are shown in Figure \ref{eqlocs}.
+Most of these earthquakes occur around the margins of the basin at locations near where we have mapped faults.
+Notably, there are three earthquakes shown in the eastern part of the Shargyn Basin.
+We lack field or satellite data constraining the location of any fault through the basin here, so I did not map a fault.
+Furthermore, the error on some of the earthquake locations is large.

### 2013-05-09 writing/paper/abstract.tex
almost done

@@
 The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources.
-We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
+We examine the fault kinematics surrounding the 150 km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system.
 The studies were performed using satellite data and targeted field transects.

### 2013-05-09 writing/paper/background.tex
almost done

@@
 Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models.
-The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citet{Tapponnier1982}\citep{Peltzer1996}.
-This hypthesis suggests that most motion has been accomodated by large strike-slip and thrust faults.
+The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citep{Tapponnier1982}\citep{Peltzer1996}.
+This hypothesis suggests that most motion has been accomodated by large strike-slip and thrust faults.
 The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993}\citep{England1997}.
-In these models, the important effects continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.
+In these models, the important effects are continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.
 
@@
 The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982}\citep{Peltzer1996}\citep{Houseman1993}.
-But, because of its orientation, active EASS rollback would exert stresses in a similar orientation to the Indo-Eurasian collision.
+But, because of its orientation, active EASS rollback would exert similarly oriented stresses to the Indo-Eurasian collision.
 However, the potential for far-field stresses from subduction zone rollback is not well constrained.
@@
 Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995}\citep{Schellart2005a}.
-A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone.
+A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000 km from the subduction zone.
 Additional constraints on the active and past deformation in western Mongolia could influence these broad theories of Asian deformation.
@@
 For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high.
-GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
+GPS measurements (Figure \ref{broad}) show that ongoing motion is toward the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}.
 The average rate of motion is approximately 5 mm/year relative to a stable Eurasia as defined by \citet{Calais2003}.
@@
 The geologic slip rate for a major transpressional fault, the Hoh-Serh-Tsagaan-Salaa Fault, is ~0.3 mm/yr of shortening and ~0.9 mm/yr of dextral slip.
-This consitutes ~20\% of the total strain accumulation in the Altai.
+This constitutes ~20\% of the total strain accumulation in the Altai.
 \citep{Frankel2010}
@@
 North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures.
-The deformation is diffuse over a 250-350km north-south extent.
+The deformation is diffuse over a 250-350 km north-south extent.
 The range accomodate the eastwards and northeastwards deformation as measured by GPS.
-At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}.
+At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan.
+Some estimates suggest 30-40 km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}.
 There are no slip estimates for the Bogd Fault System in the north.
-However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.
+However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10 km of slip \citep{Cowie1992}.
 Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.
@@
 	Previous researchers have suggested that the Hangay Dome in central Mongolia (Figure \ref{broad}) is a rigid block causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems) \citep{Cunningham2005a}.
-However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome.
+However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}.
 This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south, and may not be as rigid as previously thought\citep{Walker2006}\citep{Walker2007}\citep{Walker2008}.
@@
 	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system.
-It is approximately 150km in east-west extent and 100km in north-south extent.
+It is approximately 150 km in east-west extent and 100km in north-south extent.
 A small ephemeral lake is located in the center of the basin.
-The basin margin varies between 1500m and 2000m, while the center is at an elevation of 963m.
+The basin margin varies between 1500 meters and 2000 meters, while the center is at an elevation of 963 meters.
 The basin is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}.
 To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there.
+The Khantaishir Range to east has been the subject of petrological studies concerning the ophiolite located there.
 However, the mapping from these studies has not seriously examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.
@@
 Because strike-slip motion on both faults must be transferred into uplift, an uplifted wedge of material forms between the two faults.
+Shown in Figure \ref{basintypes}E
 \end{itemize}
@@
 The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
-Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
+On the northern boundary of the Shargyn Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.
 
@@
 
+The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
+The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
+There is a close correspondence between the major faults and this structural fabric.
+
 \begin{figure}[h!]
@@
 The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny.
-It is not shaded because it is probably not so mechanically rigid.
+The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Central Asian Orogeny.
+It is not shaded because it is probably not mechanically rigid.
 Modified after \citet{Cunningham2005a}}
@@
 \end{figure}
-
-The structural grain of these Paleozoic orogenies could affect the location of present day faults, because the metamorphic foliation represents a major plane of weakness.
-The dominant structural grain of the Central Asian Orogeny is shown in Figure \ref{grain}.
-There is a close correspondence between the major faults and this structural fabric.
-
 \clearpage

### 2013-05-09 writing/paper/discussion.tex
almost done

@@
   \caption{Schematic fault map of the Shargyn Basin region.
-The extent is approximately the same as the extent in Figure \ref{regional}.
+The extent is approximately the same as the extent in Figure \ref{intro}.
 Description of the structure is in the text.
@@
 The dotted lines represent the dominant foliation strike.
-}
+The inset in the upper right is an even simpler description of the interacting stepover and wedge uplift.}
   \label{schematic}
@@
 With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
+Dragging can be clearly seen in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
 However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault and dragging is not seen.

### 2013-05-09 writing/paper/intro.tex
removed sections on gravity
almost done

@@
 	In this paper, I examine the intersection of the Altai Range and the Gobi-Altai Range.
-The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia.
+The Altai Range and the Gobi-Altai Range are two large, actively forming mountain ranges in central Asia.
 The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}.
@@
 
-	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
+	The 150 km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
 Here, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
-The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays.
-I also use gravity data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.
+The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection, and termination thrust splays.
+I also discuss the active faulting in the context of pre-existing foliation and structural weaknesses.

### 2013-05-09 writing/paper/structure.tex
almost done

@@
 \subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region.
+During the summer of 2012, Claire Bucholz, myself, and Yeson Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region.
 We spent part of the time focusing on the active faults.
-We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
+We performed 10-20 km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
 We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
@@
   \centering
-  \includegraphics[width = 7in, height = 9.42in]{khantaishir_joint.png}
+  \includegraphics[width = 6in, height = 8.07in]{khantaishir_joint.png}
   \caption{A northeast-southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts paired with a Landsat image.
@@
 This zone is shown in Figure \ref{faultzones}A.
-Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$.
+Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$^{\circ}$ and 320$^{\circ}$.
 Fracture dip was similar to the dip of the UKT itself.
@@
   \centering
-  \includegraphics[width = 4in, height = 6in]{faultscollage.png}
+  \includegraphics[width = 4in, height = 4.53in]{faultscollage.png}
   \caption{Photographs of exposed faults.
 Field workers are used as scale bars.
-A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'.
-B) An outcrop of the exposed thrust in cross-section G-G'}
+A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section A-A'.
+B) An outcrop of the exposed thrust in cross-section X1-X1'}
   \label{faultzones}
@@
 Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the time-period before the LKT was active.
-The deeply drainages between the LKT and the UKT also indicate that the LKT is a later uplift.
+The deep drainages between the LKT and the UKT also indicate that the LKT is a later uplift.
 A major drainage cuts through the uplift adjacent to the LKT.
@@
 
-The Butara Thrust is exposed 25km west of cross-section B-B'.
+The Butara Thrust is exposed 25 km west of cross-section B-B'.
 Data from a transect performed at this location is shown in cross-section X1-X1' in Appendix A.
-Based on the outcrop, the fault strikes E-W and dips 15$\circ$.
+Based on the outcrop, the fault strikes E-W and dips 15$^{\circ}$.
 The foot wall consists of low grade metamorphosed mudstones and sandstone.
@@
   \centering
-  \includegraphics[width = 7in, height = 8.88in]{intersection_wedge_joint.png}
+  \includegraphics[width = 6in, height = 7.68in]{intersection_wedge_joint.png}
   \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
@@
 Slip sense is shown with thrust teeth or lateral slip arrows.
+The Main Mongolian Lineament is marked by the dark black line -- dashed red and black where it overlaps with a fault.
 Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.
@@
 At the eastern corner of the intersection wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking.
-At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section C-C'.
+At the location of this bend, there is a small (2 km wide) transtensional basin, shown in cross-section C-C'.
 This has been previously identified on satellite imagery by \citet{Cunningham2010}.
@@
 The small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a higher local degree of damage to the bedrock.
+In addition, there was a high degree of fracturing.
 
@@
   \includegraphics[width = 4in, height = 2.0in]{shargyn_fault_streams.png}
-  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
+  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section F-F'.}
   \label{shargynoffsetstreams}

### 2013-05-10 writing/paper/conclusion.tex
done with presentation

@@
 Further uplifts surround the basin due to termination thrust splays from the various strike-slip faults.
-Thus, the structure Shargyn Basin region is a prime example of the features that are common in intraplate transpressional orogenies and the part that pre-existing structural weaknesses can play in defining their locations.
+In addition, the locations of these faults closely follow the local foliation, bedding, or prior faulting planes.
+Thus, the structure Shargyn Basin region is a prime example of the features that occur in intraplate transpressional orogenies and the part that pre-existing structural weaknesses can play in defining their locations.

### 2013-05-10 writing/paper/discussion.tex
done with presentation

@@
-\subsection{Interpretation and Discussion}
+\section{Interpretation and Discussion}
 	The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.
@@
 One answer to this questions involves pre-existing weaknesses in the crust.
-Near the stepover, the Bogd Fault bends southward 20-30 degrees.
-For most of its length, it closely follows the Main Mongolian Lineament.
-The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
-The bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
+For most of its length, the Bogd Fault closely follows the Main Mongolian Lineament.
+he Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
+Near the stepover on the east side of the basin, the Bogd Fault bends southwards 20-30 degrees.
+This bend in the Bogd Fault may be due to a bend in the Main Mongolian Lineament.
 \citet{Windley2007} has mapped the Main Mongolian Lineament as bending southwards in this area as well.
+The trace of the Main Mongolian Lineament through the intersection wedge is shown in \ref{intersection_wedge_joint}.
 On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.

### 2013-05-10 writing/paper/main.tex
sending it to oli
removed sections on gravity
finalizing
done with presentation

@@
 \usepackage{amssymb}
-\usepackage[demo]{graphicx}
+\usepackage{graphicx}
 \usepackage{natbib}
@@
 \degreeyear{2013}
-\copyrightnoticetext{\copyright~ 2013 Thomas Ben Thompson \newline
-Anyone is permitted to reproduce and distribute copies of thesis so long as they give me credit.}
+\copyrightnotice{Thomas Ben Thompson}
 \thesisdate{May 10, 2013}
 \supervisor{Oliver Jagoutz}{Assistant Professor}
-\chairman{Samuel Bowring}{Chair, Committee on Undergraduate Program}
+\chairman{Richard P. Binzel}{Chair, Committee on Undergraduate Program}
 \maketitle
@@
 I thank the UROP office and the EAPS department for supporting my field work in Mongolia last summer.
+I also thank Yeson Erdene and Agi.
 
@@
 \include{structure}
-\include{geophysical}
-\include{basin}
+\include{discussion}
 \include{conclusion}
+\appendix
+\include{xsecappendix}
 \bibliographystyle{plainnat}

### 2013-10-07 writing/jaes_article/abstract.tex
starting jaes article

--- 2013-05-09 writing/paper/abstract.tex
@@
-\begin{abstractpage}
+\begin{abstract}
 Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision.
@@
 This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
-\end{abstractpage}
+
+Keywords: Intraplate Faulting, Basin Kinematics, Transpressional Strike-slip, Altai Fault System, Gobi-Altai Fault System
+\end{abstract}

### 2013-10-07 writing/jaes_article/background.tex
starting jaes article

--- 2013-05-09 writing/paper/background.tex
@@
 	\centering
-	\includegraphics[width = 7in, height = 6in]{asia.png}
+	\includegraphics[width = 6in, height = 5.1in]{asia.png}
 	\caption{"Simplified tectonic map of East Asia and the Western Pacific illustrating latest Cretaceous and Cenozoic faults in East Asia.
@@
 	\centering
-	\includegraphics[width = 7.0in, height = 4.09in]{broad.png}
+	\includegraphics[width = 6.0in, height = 3.40in]{broad.png}
 	\caption{Shaded relief map of Mongolia with fault overlays and seismic moment tensor solutions for recorded earthquakes.
@@
 	\centering
-	\includegraphics[width = 7in, height = 4.2in]{intro.png}
+	\includegraphics[width = 6in, height = 3.6in]{intro.png}
 	\caption{Landsat Imagery basemap showing important locations for this study.
@@
 The red lines are labeled cross-sections that will be discussed in the Structural Evidence section.
-A GPS vector near the city of Altai is shown in the northeast.
-Earthquake focal mechanisms in the area are also shown.}
+A GPS vector near the city of Altai is shown in the northeast.}
 	\label{intro}
@@
   \caption{Schematic map of the dominant structural grain in western Mongolia.
+The dark gray lines show the strike of the structural grain.
 The study area is boxed for reference.

### 2013-10-07 writing/jaes_article/discussion.tex
starting jaes article

--- 2013-05-10 writing/paper/discussion.tex
@@
 For most of its length, the Bogd Fault closely follows the Main Mongolian Lineament.
-he Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side.
+The Main Mongolian Lineament is probably a suture zone and separates a Late Proterozoic and Early Paleozoic orogeny on the north side from a Late Paleozoic orogeny on the south side \citep{Windley2007}.
 Near the stepover on the east side of the basin, the Bogd Fault bends southwards 20-30 degrees.

### 2013-10-07 writing/jaes_article/interpretation.tex
starting jaes article

--- 2013-05-08 writing/paper/interpretation.tex
@@
-\section{Structural Interpretation}
+\section{Interpretation}
 	The Shargyn Basin is a combination of a conjugate strike slip fault intersection and a left-lateral compressional strike-slip stepover.

### 2013-10-07 writing/jaes_article/intro.tex
starting jaes article

--- 2013-05-09 writing/paper/intro.tex
@@
-\section{Motivation}
+\section{Introduction}
 	Continental deformation can occur far from plate boundaries.
@@
 
-	In this paper, I examine the intersection of the Altai Range and the Gobi-Altai Range.
+	In this paper, we examine the intersection of the Altai Range and the Gobi-Altai Range.
 The Altai Range and the Gobi-Altai Range are two large, actively forming mountain ranges in central Asia.
@@
 	The 150 km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system.
-Here, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
+Here, we describe the Shargyn Basin as the result of transpressional strike-slip kinematics.
 The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection, and termination thrust splays.
-I also discuss the active faulting in the context of pre-existing foliation and structural weaknesses.
+We also discuss the active faulting in the context of pre-existing foliation and structural weaknesses.

### 2013-10-07 writing/jaes_article/main.tex
starting jaes article

--- 2013-05-10 writing/paper/main.tex
@@
-\documentclass[12pt, doublepage, vi]{mitthesis}
+\documentclass[12pt]{article}
 \usepackage[latin1]{inputenc}
@@
 \usepackage{natbib}
-\usepackage[margin=0.75in]{geometry}
+\usepackage[margin=1.0in]{geometry}
 \usepackage{setspace}
-\renewcommand\thechapter{}
-\renewcommand\thesection{\arabic{section}}
-\renewcommand\thesubsection{\thesection.\arabic{subsection}}
+\usepackage{caption}
+\doublespacing
+\AtBeginCaption{\doublespacing}
 \begin{document}
 \title{Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.}
-\author{Thomas Ben Thompson}
-\department{Department of Earth, Atmospheric and Planetary Science}
-\degree{Bachelor of Science in Earth, Atmospheric and Planetary Sciences}
-\degreemonth{June}
-\degreeyear{2013}
-\copyrightnotice{Thomas Ben Thompson}
-\thesisdate{May 10, 2013}
-\supervisor{Oliver Jagoutz}{Assistant Professor}
-\chairman{Richard P. Binzel}{Chair, Committee on Undergraduate Program}
+\author{%
+    Thomas Ben Thompson (Corresponding Author)\\
+    Department of Earth, Atmospheric and Planetary Sciences at MIT\\
+    \texttt{tthompson@fas.harvard.edu}\vspace{40pt} \\
+    Oliver Jagoutz \\
+    Department of Earth, Atmospheric and Planetary Sciences at MIT \\
+    \texttt{jagoutz@mit.edu} \vspace{40pt} \\
+    Claire Bucholz \\ 
+    Department of Earth, Atmospheric and Planetary Sciences at MIT \\
+    \texttt{cbucholz@mit.edu} \\
+    }
 \maketitle
+
 \include{abstract}
-
-\section*{Acknowledgments}
-Primarily, I thank Claire Bucholz (EAPS PhD candidate) for wandering the Mongolian steppes and mountains and teaching me endlessly about geology and EAPS Professor Oliver Jagoutz for supervising and helping me through the whole thesis process.
-I thank the UROP office and the EAPS department for supporting my field work in Mongolia last summer.
-I also thank Yeson Erdene and Agi.
 
@@
 \include{xsecappendix}
+\section*{Acknowledgments}
+Many thanks to Yeson Erdene, a student at MUST, and Agi, our friend and driver. 
 \bibliographystyle{plainnat}

### 2013-10-07 writing/jaes_article/structure.tex
starting jaes article

--- 2013-05-09 writing/paper/structure.tex
@@
   \centering
-  \includegraphics[width = 7in, height = 5.0in]{fault_map.png}
+  \includegraphics[width = 6in, height = 4.2in]{fault_map.png}
   \caption{A map of the Shargyn Basin showing the studied faults and important locations. Slip is indicated by thrust teeth or lateral slip arrows.}
@@
   \centering
-  \includegraphics[width = 7in, height = 6.83in]{haliun_joint.png}
+  \includegraphics[width = 6in, height = 5.83in]{haliun_joint.png}
   \caption{A cross-section through the Haliun Thrust and Butara Thrust paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht!]
   \centering
-  \includegraphics[width = 6in, height = 7.68in]{intersection_wedge_joint.png}
+  \includegraphics[width = 5in, height = 6.68in]{intersection_wedge_joint.png}
   \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
@@
   \centering
-  \includegraphics[width = 7in, height = 7.5in]{lil_dariv_joint.png}
+  \includegraphics[width = 6in, height = 6.6in]{lil_dariv_joint.png}
   \caption{A cross-section through the Little Dariv Mountain area paired with a Landsat image of the surrounding area. Important locations and lithology are noted on the satellite image.}
@@
   \centering
-  \includegraphics[width = 7in, height = 6.85in]{neshargyn_joint.png}
+  \includegraphics[width = 6in, height = 5.85in]{neshargyn_joint.png}
   \caption{A cross-section through the Shargyn Fault in the northeast corner of the Shargyn Basin paired with a Landsat image of the surrounding area.

### 2013-10-18 writing/jaes_article/background.tex
working on jaes article

@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
 	\centering
@@
 
-	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}\citep{Walker2007}\citep{Cunningham2005a}\citep{England1997}.
+	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010, Walker2007, Cunningham2005a, England1997}.
 Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models.
-The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citep{Tapponnier1982}\citep{Peltzer1996}.
+The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citep{Tapponnier1982, Peltzer1996}.
 This hypothesis suggests that most motion has been accomodated by large strike-slip and thrust faults.
-The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993}\citep{England1997}.
+The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993, England1997}.
 In these models, the important effects are continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.
@@
 	Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}.
-The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982}\citep{Peltzer1996}\citep{Houseman1993}.
+The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982, Peltzer1996, Houseman1993}.
 But, because of its orientation, active EASS rollback would exert similarly oriented stresses to the Indo-Eurasian collision.
@@
 Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins.
-Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995}\citep{Schellart2005a}.
+Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995, Schellart2005a}.
 A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000 km from the subduction zone.
@@
 In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia.
-The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}.
+The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998, Okal1976a, Pollitz2003}.
 The Bogd Earthquake has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.
@@
 Many of these faults terminate in large thrust systems like the Tian Shan.
-One explanation invokes bookshelf-like rotation along a long left-lateral shear zone creates the right-lateral faulting.
+One explanation suggests bookshelf-like rotation along a long left-lateral shear zone creates the right-lateral faulting.
 \citep{Davy1988}\citep{Bayasgalan2005a}  This model implies that the development of the Altai and Gobi-Altai is intimately linked to that of the Tian Shan.
@@
 The region is dominated by NW-SE striking right-lateral slip.
-The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults.
+The primary features are strike-slip faults, but much of the topography is created by thrusts oriented ~30 degrees off the main strike slip faults.
 Others are linked by a right-slip faults.
@@
 
-\subsection{Shargyn Basin}
+\subsection{The Shargyn Basin}
 
@@
 To the east and south, the tectonic regime is less certain.
-The Khantaishir Range to east has been the subject of petrological studies concerning the ophiolite located there.
-However, the mapping from these studies has not seriously examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.
+The Khantaishir Range to east has been the subject of petrological studies concerning the ophiolite located there \citep{Yarmolyuk2011}.
+However, the mapping from these studies has not closely examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.
 The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, are also poorly understood.
 
-\begin{figure}[h!]
+\begin{figure}[p!]
 	\centering
@@
 	\caption{Landsat Imagery basemap showing important locations for this study.
-The large white box surrounding the Shargyn Basin is the new mapping from this study.
-The white area surrounding the Dariv Range indicates mapping performed by \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.
+The Dariv Range has been mapped \citet{Dijkstra2006} and as a part of unpublished research by Claire Bucholz.
 The area surrounding the Sutai Range was mapped by \citet{Cunningham2003}.
-The small white boxes around the Shargyn Basin show areas that we studied in the field.
-The red lines are labeled cross-sections that will be discussed in the Structural Evidence section.
+The black lines are labeled cross-sections that will be discussed in the Structural Evidence section.
 A GPS vector near the city of Altai is shown in the northeast.}
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 4.5in, height = 5.67in]{basins.png}
-  \caption{Schematic diagrams showing faulting and deposition for common transpressional uplifts and basins.}
+  \includegraphics[width = 5.0in, height = 4.15in]{basins.png}
+  \caption{Schematic diagrams showing faulting and deposition for relevant types of transpressional uplifts and basins.}
   \label{basintypes}
@@
 
-Many different types of basins form in strike-slip and compressional settings.
-Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}.
-Some of the settings we might expect:
-\begin{itemize}
-	\item Foreland basins form in front of a simple thrust system. Flexural subsidence can result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
-	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
-	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure.
-Shown in Figure \ref{basintypes}C
-	\item Compresional stepover basins form similarly to restraining bend basins.
-However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault.
-In the compressional case, the basin is essentially a foreland basin in front of the thrust.
-Arguably, these fault systems are just large restraining bends.
-Shown in Figure \ref{basintypes}D.
-	\item An intersection wedge can form at a conjugate strike-slip intersection.
-Because strike-slip motion on both faults must be transferred into uplift, an uplifted wedge of material forms between the two faults.
-Shown in Figure \ref{basintypes}E
-\end{itemize}
+A variety of features form in strike-slip and compressional settings.
+Futhermore, complex interactions can occur at the intersection of fault systems \citep{Busby1995}.
+Some of the basic building blocks in such a setting: Simple thrusts can cause flexural subsidence and local foreland basins.
+Piggyback basins can form structurally above a thrust.
+The adjacent uplift provide a sediment source.
+Uplifts can form at strike-slip stepovers and bends.
+Often such stepovers form positive flower fault structures \citep{Cunningham2005a}.
+Significant deposition occurs adjacent to these uplifts.
+Finally and importantly for the Shargyn Basin, at the intersection of conjugate strike-slip systems, an uplifted wedge of material can form.
+This wedge forms because motion on both faults must terminate in uplift.
+The most important of these features are shown in Figure \ref{basintypes}.
 
-	A goal of this paper is to describe the Shargyn Basin in these terms.
-A summary of the results is in the Structural Interpretation section.
+	The goal of this paper is to describe the Shargyn Basins in terms of such simple kinematic building blocks.
 Nearby regions have already been investigated.
@@
 It represents the southeastern-most uplift in the Altai.
-The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003}\citep{Howard2006}.
+The Sutai Range is an active restraining bend in the Altai fault system\citep{Cunningham2003, Howard2006}.
 A stepover in the right lateral fault has resulted in progressively more thrusting.
@@
 	To the east of the Sutai Range, the Dariv Thrust uplifts the Dariv Mountains.
-\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust.
+\citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of a block that has been uplifted by the Dariv Thrust.
 The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range.
@@
 
-The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different ages.
+The Shargyn Basin lies near the Main Mongolian Lineament (MML), a boundary between rocks of very different ages \citep{Badarch2002}.
 North of the MML, Proterozoic and early Paleozoic units are present, whereas on the southern side of the MML, there are Silurian and Devonian gneiss, schist and felsic igneous rocks.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6.4in, height = 4.5in]{bedrockfigure.png}
-  \caption{Schematic tectonostratigraphic map of Mongolia.
+  \includegraphics[width = 6in, height = 4.5in]{bedrockfigure.png}
+  \caption{Schematic tectonostratigraphic map of the Central Asian Orogenic Belt in Mongolia.
 The area we studied is located in the outlined box.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
@@
   \caption{Schematic map of the dominant structural grain in western Mongolia.
-The dark gray lines show the strike of the structural grain.
+The black lines show the strike of the structural grain.
 The study area is boxed for reference.
@@
 The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.
-The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Central Asian Orogeny.
+The Hangay Dome consists of microcontinental fragments that accreted during the early stages of the Central Asian Orogeny.
 It is not shaded because it is probably not mechanically rigid.

### 2013-10-18 writing/jaes_article/discussion.tex
working on jaes article

@@
 However, some motion remains on the continuation of the Bogd Fault.
-The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that I will call the intersection wedge.
+The remainder of this motion is transferred into multiple thrust splays, the last of which combines with terminal thrust splays from the Tonhil Fault to form an uplifted region that we will call the intersection wedge.
 Similar thrust splays on the Shargyn Fault uplift the Dariv Range and link with thrust faults in the Dariv Basin and the Sutai Range.
@@
 Description of the structure is in the text.
-The shaded areas are basins.
-Unshaded areas are mountainous.
+The shaded areas are depositional (basins).
+Unshaded areas are uplifted.
 The dotted lines represent the dominant foliation strike.
@@
 
-	Thus, the eastern part of the Shargyn Basin is a stepover basin forming from the Khantaishir Thrusts.
-The western and southern part of the basin is a foreland basin from the uplifted intersection wedge and the many termination thrust splays of the Bogd Fault.
-On the northern boundary, the uplift of the Dariv Range is concentrated on the Dariv Thrust and the Shargyn Fault separates the low-lying basin from the adjacent mountain range while not accomodating the uplift.
-
-	Notably, the kinematics for the Shargyn Basin explain a large basin without any extension.
+	Notably, these kinematics for the Shargyn Basin explain a large basin without any extension.
 Despite being deeper than most basins in the area (963m above sea level at the deepest), it would unexpected for the Shargyn Basin to have formed by any extensional process.
@@
 The small transtensional basins near the intersection wedge were the only extensional features we found.
-However, these transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
+These transtensional basins are multiple orders of magnitude smaller than the Shargyn Basin.
 Further to the northeast, there are small normal faults linked by a major strike-slip fault \citep{Walker2006}.
@@
 	With a kinematic description of the Shargyn Basin in hand, it is interesting to ask why such structures formed in their current locations.
+Nearly all the study areas have foliations striking parallel or subparallel to the most prominent nearby fault.
+Hence, the foliations fit well with a model where faults form along structural weaknesses.
+The foliations could also be explained by dragging along the fault.
+With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
+Dragging can be clearly seen in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
+However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault.
+Dragging is not seen in these areas.
+So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.
+
         One answer to this questions involves pre-existing weaknesses in the crust.
@@
 On a broader scale, the Gobi-Altai and Altai ranges both follow the approximate structural grain in Figure \ref{grain}.
-
-	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes parallel or close to parallel to the important fault in the region.
-The foliations fit well with a model where faults form along structural weaknesses.
-The foliations could also be explained by dragging along the fault.
-With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike.
-Dragging can be clearly seen in the satellite imagery of the beds along the Bogd Fault in the intersection wedge.
-However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault and dragging is not seen.
-So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.

### 2013-10-18 writing/jaes_article/structure.tex
working on jaes article

@@
 Previous studies have examined Mongolian active tectonics from a remote sensing perspective \citep{Cunningham2005a}\citep{Cunningham2010}\citep{Tapponnier1979}\citep{Walker2007}.
-I performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
+We performed a more detailed remote sensing survey of the Shargyn Basin region using Landsat 7 TM+ satellite imagery.
 Surface features indicating active motion were identified.
 Because the region is arid, many important features are clearly visible.
-For example, the trace of Tonhil Fault on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
+For example, the trace of Tonhil Fault (see Figure \ref{faultmap}) on the western margin of the Shargyn Basin is visible as a very sharp linear topographic feature.
 Using such features from Landsat imagery was useful in identifying important features for further investigation in the field.
@@
 
-In the remote sensing studies, I focused on identifying prominent faults using steep topography, sharp color changes (lithology changes), fault scarps, and offset features.
-I also mapped clearly distinguishable lithologic units using color, weathering, and foliation changes visible in the Landsat imagery.
-Where possible, I noted the strike of the foliation.
+In the remote sensing studies, we focused on identifying prominent faults using steep topography, sharp color changes (lithology changes), fault scarps, and offset features.
+We also mapped clearly distinguishable lithologic units using color, weathering, and foliation changes visible in the Landsat imagery.
+Where possible, we noted the strike of the foliation.
 Foliation is useful to map out the Paleozoic structural grain in the region.
-I also noted the location and distribution of alluvial fans.
-Locations where alluvial fans are forming above active thrusts are important because that might indicate a temporal or slip difference between multiple adjacent faults.
+We also noted the location and distribution of alluvial fans.
+Locations where alluvial fans are forming above active thrusts are important because such fans might indicate a temporal or slip difference between multiple adjacent faults.
 
 \subsection{Field Mapping}
-During the summer of 2012, Claire Bucholz, myself, and Yeson Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region.
-We spent part of the time focusing on the active faults.
-We performed 10-20 km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}).
+The field work was performed during the summer of 2012.
+We performed 10-20 km transects through areas with important faults, as previously identified by satellite image.
 We mapped at a broad scale, but made careful note of features indicating recent brittle deformation.
@@
 Published mapping in the Sutai Range by \citet{Cunningham2003} and in the Dariv Range by \citet{Dijkstra2006} is incorporated.
-Further unpublished mapping of the Dariv Range by Claire Bucholz is incorporated.
+Further unpublished mapping of the Dariv Range by (Bucholz and Jagoutz) is incorporated.
 We performed new mapping surrounding the Shargyn Basin.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6in, height = 4.2in]{fault_map.png}
-  \caption{A map of the Shargyn Basin showing the studied faults and important locations. Slip is indicated by thrust teeth or lateral slip arrows.}
+  \includegraphics[width = 6in, height = 3.46in]{fault_map.png}
+  \caption{A map of the Shargyn Basin showing the studied faults and important locations. Slip is indicated by thrust teeth or lateral slip arrows. Question marks indicate unknown slip sense.}
   \label{faultmap}
@@
 
-\subsection{Khantaishir Thrusts}
+\subsection{Khantaishir Stepover and Thrusts}
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6in, height = 8.07in]{khantaishir_joint.png}
+  \includegraphics[width = 6in]{khantaishir_joint.png}
   \caption{A northeast-southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts paired with a Landsat image.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 4in, height = 4.53in]{faultscollage.png}
+  \includegraphics[width = 4in]{faultscollage.png}
   \caption{Photographs of exposed faults.
@@
 A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section A-A'.
-B) An outcrop of the exposed thrust in cross-section X1-X1'}
+B) An outcrop of the Upper Haliun Thrust in cross-section X1-X1'}
   \label{faultzones}
@@
 The trace of the Lower Khantaishir Thrust (LKT) is located 10km to the southwest of the UKT.
-South from the UKT, there is a transition from recent alluvium to more lithified, potentially older alluvium.
-Then, for 2 kilometers before reaching the LKT, there is a thick stack of sandstone and basalt flows, with interlayered sandy carbonate beds.
-Finally, there is a very sharp topographic transition and lithologic shift to recent alluvium.
-This marks the fault trace for the LKT.
-The sedimentary sequence northeast of the LKT indicates the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area.
+South from the UKT, there is a transition from recent alluvium to more compacted, potentially older alluvium.
+Within two kilometers of the LKT, there is an uplifted thick stack of sandstone and basalt flows, with interlayered sandy carbonate beds.
+Marking the fault trace of the LKT is a very sharp topographic transition and lithologic shift to recent alluvium.
+The sedimentary sequence northeast of the LKT might indicate the presence of a minor basin in the area at a point after the widespread metamorphism affecting the area.
 
@@
 Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the time-period before the LKT was active.
-The deep drainages between the LKT and the UKT also indicate that the LKT is a later uplift.
+The deep drainages between the LKT and the UKT may also indicate that the LKT is a later uplift.
 A major drainage cuts through the uplift adjacent to the LKT.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6in, height = 5.83in]{haliun_joint.png}
-  \caption{A cross-section through the Haliun Thrust and Butara Thrust paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
+  \includegraphics[width = 6in]{haliun_joint.png}
+  \caption{A cross-section through the Lower Haliun Thrust and Upper Haliun Thrust paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
   \label{haliun_joint}
@@
 
-Along the southern margin of the basin, we performed two transects across two thrusts splayed off the Bogd left-lateral strike slip fault.
-We will refer to the northern thrust as the Haliun Thrust and the southern thrust as the Butara Thrust.
-Above (south) of the Haliun Thrust, there is a complex sequence of metasedimentary and metavolcanics rocks that are intruded by a large gabbro body.
-There might be a fault-adjacent anticline.
-I do not know whether such an anticline is actively forming or a remnant from previous orogenies.
-The foliation in the fault-adjacent units is parallel to the Haliun Thrust and dips to the south.
+Along the southern margin of the basin, we performed a pair of transects across two thrusts splayed off the Bogd left-lateral strike slip fault (see Figure \ref{haliun_joint}).
+We will refer to the northern thrust as the Lower Haliun Thrust (LHT) and the southern thrust as the Upper Haliun Thrust (UHT).
+Structurally above (south) of the LHT, there is a complex sequence of metasedimentary and metavolcanics rocks that are intruded by a large gabbro body.
+The foliation in the fault-adjacent units is parallel to the LHT and dips to the south.
 Along the basin margin, elevations decrease to the west.
-The decrease in elevation indicates the uplift tapers off further from the Bogd strike-slip fault.
+The decrease in elevation suggests the uplift tapers off further from the Bogd strike-slip fault.
 
-To the south of the Haliun Thrust, the Butara Thrust uplifts 3000m high peaks in the center of the Bogd Range.
-To the north of the Butara Thrust, we observed the same metasedimentary sequence as is adjacent to the Haliun Thrust.
-To the south of the Butara Thrust, there are sandstones and basaltic units.
-Just to the west of the cross-section, alluvium is accumulating south of the Haliun Thrust, indicating that more of the uplift is occuring on the Butara Thrust.
-To the west of cross-section, the high topography is almost entirely south of the Butara Thrust.
+South of the LHT, the UHT uplifts 3000m high peaks in the center of the Bogd Range.
+North of the UHT, we observed the same metasedimentary sequence as is adjacent to the LHT.
+South of the UHT, there are sandstones and basaltic units.
+Just to the west of the cross-section, alluvium is accumulating structurally above the LHT, indicating that substantial uplift is occuring on the UHT.
+West of the cross-section, the highest topography is almost entirely south of the UHT, reinforcing the idea that uplift is concentrated on the UHT.
 
-The Butara Thrust is exposed 25 km west of cross-section B-B'.
+The UHT is exposed 25 km west of cross-section B-B'.
 Data from a transect performed at this location is shown in cross-section X1-X1' in Appendix A.
@@
 The hanging wall consists of greenschist.
-Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge.
+Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of gouge.
 Kinematic indicators suggest thrust sense of motion.
+The topography rises sharply directly south of the location of the fault.
 A photograph is shown in Figure \ref{faultzones}B.
-The topography rises sharply directly south of the location of the fault.
 
 Further west, there is a third thrust splay off the Bogd Fault that can be seen in Figure \ref{faultmap}.
-But, we did not visit this fault in the field map.
-The satellite imagery suggests that it is similar in nature to the Haliun Thrust and the Butara Thrust.
+We did not visit this fault in the field.
+The satellite imagery suggests that it is similar in character to the LHT and the UHT.
 
@@
   \centering
-  \includegraphics[width = 5in, height = 6.68in]{intersection_wedge_joint.png}
+  \includegraphics[width = 5in]{intersection_wedge_joint.png}
   \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin.
-Mapped active faults are shown by solid red lines.
+Mapped active faults are shown by solid black lines.
 Slip sense is shown with thrust teeth or lateral slip arrows.
-The Main Mongolian Lineament is marked by the dark black line -- dashed red and black where it overlaps with a fault.
+The Main Mongolian Lineament is marked by the dashed white line -- dashed white and black where it overlaps with a fault.
 Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge.
@@
 
-At cross-section D-D' (shown in Figure \ref{intersection_wedge_joint}, we performed a transect through the series of thrusts separating the basin from the intersection wedge.
+At cross-section D-D' (shown in Figure \ref{intersection_wedge_joint}), we performed a transect through the series of thrusts separating the basin from the intersection wedge.
 At the site of the transect, there are three thrusts.
-The lowest thrust marks the basin margin.
+The structurally lowest thrust marks the basin margin.
 To the southwest of this thrust, there is a large hornblende-rich diorite intrusion with minor intrusions of granite included.
-Close to the middle thrust, there is sedimentary sequence that includes an actively mined coal bed.
+Close to the middle thrust, there is organic-rich sedimentary sequence including an actively mined coal bed.
 The middle thrust causes a noticeable topographic jump.
@@
 This is interpreted as the furthest northern extent of the southwestern-most of the intersection wedge thrusts.
-For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss and schist from the metasedimentary rocks, indicating that it is an important structural boundary.
+For most of its length, satellite imagery of this fault shows some topographic expression and where observed in the field, it always separates the gneiss from the lower-grade metasedimentary rocks.
+This indicates that it is an important structural boundary.
 
@@
 The small depositional region and the topographic rise would be unlikely without recent motion.
-In addition, there was a high degree of fracturing.
+In addition, there was a high degree of localized fracturing.
 
@@
 The fault strikes WSW-ENE.
-In the cross-sections, I have assumed that the Shargyn Fault is approximately vertical because of its lateral slip and the lack of other constraints.
+In the cross-sections, we have assumed that the Shargyn Fault is approximately vertical because of its lateral slip and the lack of other constraints.
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6in, height = 6.6in]{lil_dariv_joint.png}
+  \includegraphics[width = 6in]{lil_dariv_joint.png}
   \caption{A cross-section through the Little Dariv Mountain area paired with a Landsat image of the surrounding area. Important locations and lithology are noted on the satellite image.}
@@
 \clearpage
-The western end of the Shargyn Fault, nicknamed Little Dariv Mountain, exhibits thrusts in a small positive flower structure, seen in cross-section E-E' in Figure \ref{lil_dariv_joint}.
+The western end of the Shargyn Fault, nicknamed Little Dariv Mountain \citep{Howard2006}, exhibits thrusts in a small positive flower structure, seen in cross-section E-E' in Figure \ref{lil_dariv_joint}.
 The positive flower structure has been identified by previous researchers \citep{Howard2006}.
@@
 The southern fault can be seen in the satellite imagery as an uplifted trace of bedrock amongst a sea of basin sediments.
-It may be related to a large anticline visible on the satellite imagery west of Little Dariv Mountain.
 The northern boundary fault is the inferred continuation of the main Shargyn Fault trace.
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 6in, height = 5.85in]{neshargyn_joint.png}
+  \includegraphics[width = 6in]{neshargyn_joint.png}
   \caption{A cross-section through the Shargyn Fault in the northeast corner of the Shargyn Basin paired with a Landsat image of the surrounding area.
@@
 On the north side of the fault, there are amphibolite facies gneiss and schist units with south-dipping foliations.
-There are also some large granitoid intrusions.
+There are also substantial granitoid intrusions.
 Outcrops of garnet schist confirm the higher metamorphic grade.
-To the west, offset stream beds shown in Figure \ref{shargynoffsetstreams} provide one of the clearest indications of the active slip sense.
+To the west, offset stream beds shown in Figure \ref{shargynoffsetstreams} provide one of the clearest indications of the active left-lateral slip sense.
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering
@@
 The difference in metamorphic grade between the granite and the metamorphic unit indicates the granite was intruded after metamorphism.
-However, the contact between the granite and the metamorphic is perfectly straight, supporting the contact's interpretation as a fault.
+However, the contact between the granite and the metamorphic unit is very straight and shows no features of an intrusive contact, supporting the interpretation as a fault.
 The topographic depression along the fault provides further evidence.
@@
 \citet{Bayasgalan2005a} provides a good review of the earthquake focal mechanisms in western Mongolia.
-Only a couple mechanisms are located in the Shargyn Basin.
 On the Shargyn Fault, near cross-section C-C', a focal mechanism was calculated showing exclusively left-lateral slip.
 In the northern corner of the intersection wedge, a focal mechanism was calculated either showing low-angle south directed thrusting or high-angle north directed thrusting.
-Both possibilities are reasonable given the location of the earthquake.
-
-\begin{figure}[h!]
+Both possibilities are reasonable given the location of the earthquake and the sense of slip for the mapped faults.
+\begin{figure}[p!]
   \centering
-  \includegraphics[width = 5in, height = 2.88in]{eq_locs.png}
+  \includegraphics[width = 6in]{eq_locs.png}
   \caption{Shaded relief map of the Shargyn Basin showing blue plus signs where earthquakes have occured between 1960 and 2013.
@@
 The locations for all the earthquakes available from the IRIS Data Management System  in the Shargyn Basin region are shown in Figure \ref{eqlocs}.
-Most of these earthquakes occur around the margins of the basin at locations near where we have mapped faults.
-Notably, there are three earthquakes shown in the eastern part of the Shargyn Basin.
-We lack field or satellite data constraining the location of any fault through the basin here, so I did not map a fault.
-Furthermore, the error on some of the earthquake locations is large.
+Importantly, most of these earthquakes occur around the margins of the basin at locations near where we have mapped faults.
+There are three earthquakes shown in the eastern part of the Shargyn Basin.
+Satellite imagery weakly suggests a fault passes through the area.
+However, the location error on these earthquake locations is large due to their low magnitude and significant distance to installed seismometers.
